---
title: "data exploring"
author: "Anne Baranger"
date: "2023-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(targets)
lapply(c("stringr","ggplot2","tidyr","dplyr","rworldmap","modi"),require,character.only=TRUE)
tar_load(mastif.data,store="target_data")
tar_load(meanClimate_species,store="target_gbif")
```

# plot data

```{r plotplot}
df.tree=mastif.data$df.tree

# df.niche=read.csv("output/sp_gbif_climate.csv",sep=" ")
#plot initial dataset and selection
worldmap <- sf::st_as_sf(getMap(resolution = "high"))


# plots of selected species
df.tree |> 
  filter(species %in% mastif.data$df.species.select$species) |> 
  select(species,lon,lat) |> 
  distinct() |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=species))+
  xlim(range(mastif.data$df.species[,c("minlon","maxlon")])) + 
  ylim(range(mastif.data$df.species[,c("minlat","maxlat")]))+
  theme_minimal()+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom")+
  labs(color="Species")


df.tree |> 
  filter(species %in% mastif.data$df.species.select$species) |> 
  select(species,lon,lat) |>
  filter(species=="pinusSylvestr") |> 
  distinct() |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=species))+
  xlim(range(mastif.data$df.species[,c("minlon","maxlon")])) + 
  ylim(range(mastif.data$df.species[,c("minlat","maxlat")]))+
  theme_minimal()+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom")+
  labs(color="Species")

```

# Compare species range with species niche

```{r}

df.tree.range.1<-df.tree |>
  mutate(fecEstSe=na_if(fecEstSe,0),
         fecEstMu=na_if(fecEstMu,0),
         fecStand=100*fecEstSe/fecEstMu,
         across(c("mat","tmin"),
                ~.x/10)
         ) |> 
  group_by(treeID,species,mat,map,tmin) |> 
  summarise(fecStandMean=mean(fecStand,na.rm=TRUE),
            fecEstMuMean=mean(fecEstMu,na.rm=TRUE)) |> 
  mutate(fecW=1-fecStandMean/100) |> 
  pivot_longer(cols=c("mat","tmin","map"),
               names_to = "clim",
               values_to = "clim_val")  |> 
  ungroup() |> 
  filter(!is.nan(fecEstMuMean)) |> 
  group_by(species,clim) |>
  summarise(opt=median(clim_val),
            low=weighted.quantile(clim_val,fecW,prob=0.025)[1],#,probs=0.1)[1],
            high=weighted.quantile(clim_val,fecW,prob=0.975)[1]) |> 
  pivot_wider(names_from = "clim",
              values_from = c("opt","high","low"),
              names_sep = ".") |> 
  pivot_longer(cols=-species,
               names_to = "clim",
               values_to = "clim_obs") |> 
  mutate(clim=paste0(str_split_i(clim,"\\.",2),".",str_split_i(clim,"\\.",1)))
df.tree.range <-df.tree |> 
  filter(!duplicated(treeID)) |> 
  mutate(across(c("mat","tmin"),
                ~.x/10)) |> 
  pivot_longer(cols=c("mat","tmin","map"),
               names_to = "clim",
               values_to = "clim_val") |> 
  group_by(species,clim) |>
  summarise(opt=median(clim_val),
            low=quantile(clim_val,probs=0.025)[1],#,probs=0.1)[1],
            high=quantile(clim_val,probs=0.975)[1]) |> #,probs=0.9)[1]) |> 
  pivot_wider(names_from = "clim",
              values_from = c("opt","high","low"),
              names_sep = ".") |> 
  pivot_longer(cols=-species,
               names_to = "clim",
               values_to = "clim_obs") |> 
  mutate(clim=paste0(str_split_i(clim,"\\.",2),".",str_split_i(clim,"\\.",1)))

df.niche= meanClimate_species |> 
  separate(species,into=c("genus","sp"),remove=FALSE) |> 
  mutate(species_l=species,
         species=paste0(substr(tolower(genus),1,8),substr(str_to_title(sp),1,8))) |> 
  rename_with(.cols=c("mat","map","tmin"),
              ~paste0(.x,".opt")) |> 
  select(-genus,-sp) |> 
  mutate(mat.range=mat.high-mat.low,
         map.range=map.high-map.low,
         tmin.range=tmin.high-tmin.low) |> 
  pivot_longer(cols=-c("species","species_l"),
               names_to = "clim",
               values_to = "clim_niche") 
df.tree.range |> 
  left_join(df.niche,by=c("species","clim")) |> 
  filter(!is.na(species_l)) |> 
  relocate(species_l,.after=species)|> 
  mutate(dif=100*(clim_obs-clim_niche)/clim_niche)
df.tree.range |> 
  left_join(df.niche,by=c("species","clim")) |> 
  filter(!is.na(species_l)) |> 
  relocate(species_l,.after=species) |> 
  mutate(dif=100*(clim_obs-clim_niche)/clim_niche) |> 
  ggplot(aes(dif,clim,color=species))+
  geom_point()


```


Overlay distribution of Gbif and mastif
```{r}
df.niche.short= 
  df.niche |>
  separate(clim,into = c("clim","range")) |> 
  pivot_wider(names_from = range,
                values_from = clim_niche)

for (sp in unique(df.tree$species)){
  plot=df.tree |>
    filter(species==sp) |> 
    mutate(fecEstSe=na_if(fecEstSe,0),
           fecEstMu=na_if(fecEstMu,0),
           fecStand=100*fecEstSe/fecEstMu,
           across(c("mat","tmin"),
                  ~.x/10)
           ) |> 
    group_by(treeID,species,mat,map,tmin) |> 
    summarise(fecStandMean=mean(fecStand,na.rm=TRUE),
              fecEstMuMean=mean(fecEstMu,na.rm=TRUE)) |> 
    mutate(fecW=1-fecStandMean/100) |> 
    pivot_longer(cols=c("mat","tmin","map"),
                 names_to = "clim",
                 values_to = "clim_val")  |> 
    ungroup() |> 
    left_join(df.niche.short,by=c("species","clim")) |> 
    ggplot()+
    geom_histogram(aes(x=clim_val,y=..density..,weight=fecW))+
    geom_point(aes(x=opt,y=0),color="darkred",size=1)+
    geom_segment(aes(x=low,xend=high,y=-0,yend=-0),color="darkred",size=1)+
    facet_wrap(~clim,
               # nrow = 3,
               scales = "free")+
    labs(title=sp)
  print(plot)
  rm(plot)
}
# 
#   ggplot()+
#   geom_histogram(aes(x=clim_val))+
#     geom_point(aes(x=opt,y=0),color="darkred",size=1)+
#     geom_segment(aes(x=low,xend=high,y=-0,yend=-0),color="darkred",size=1)+
#     facet_wrap(~clim,
#                # nrow = 3,
#                scales = "free")+
#     labs(title=sp)+
#   scale_y_log10()
```


```{r}
df.tree |> 
  filter(!duplicated(treeID)) |> 
  select(species,plotID) |> 
  unique() |> 
  group_by(species) |> 
  summarise(n=n())


```

Test if with KS test and simulating a triangular distribution
-> pas ouf comme méthode
```{r eval=FALSE, include=FALSE}
for (sp in unique(df.niche$species)){
  # get df for targetted species
  df=df.tree |> 
    filter(!duplicated(treeID)) |> 
    filter(species==sp) |> 
    select(plotID,species,mat,map,tmin) |> 
    unique() |> 
    mutate(across(c("mat","tmin"),
                  ~.x/10))  
    # pivot_longer(cols=c("mat","tmin","map"),
    #              names_to = "clim",
    #              values_to = "clim_val")  |> 
    # left_join(df.niche.short,by=c("species","clim")) |
  n=(df.tree |> 
    filter(!duplicated(treeID)) |> 
    filter(species==sp) |> 
    select(plotID) |> 
    unique() |> 
    count())$n
  niche=df.niche |> 
    filter(species==sp) |> 
    pivot_wider(names_from = clim,
                values_from = clim_niche)
  #simulate triangular distribution from gbif
  gbif.dist=list(map=matrix(rep(rtri(n, min = niche$map.low,  max = niche$map.high,  mode = niche$map.opt),50),nrow=n),
                 mat=matrix(rep(rtri(n, min = niche$mat.low,  max = niche$mat.high,  mode = niche$mat.opt),50),nrow=n),
                 tmin=matrix(rep(rtri(n, min = niche$tmin.low,  max = niche$tmin.high,  mode = niche$tmin.opt),50),nrow=n))
  # get data from mastif
  mastif.dist=list(map=df$map,
                   mat=df$mat,
                   tmin=df$tmin)
  suppressWarnings({
    res=list(map=sum(apply(gbif.dist$map,MARGIN = 2,function(x)ks.test(mastif.dist[["map"]],x)$p.value>0.05)),
             mat=sum(apply(gbif.dist$mat,MARGIN = 2,function(x)ks.test(mastif.dist[["mat"]],x)$p.value>0.05)),
             tmin=sum(apply(gbif.dist$tmin,MARGIN = 2,function(x)ks.test(mastif.dist[["tmin"]],x)$p.value>0.05)))
  }
    )
  print(sp)
  print(res)

  rm(df,niche,n,gbif.dist,mastif.dist,res)
}

```


Test difference in quantiles and optima
```{r}
df.range=df.niche |> 
  select(species,clim,clim_niche) |> 
  filter(grepl("range",clim)) |> 
  separate(clim,into=c("var","drop")) |> 
  select(-drop) |> 
  rename(range="clim_niche")
df.tree.range.1 |> 
  left_join(df.niche,by=c("species","clim")) |> 
  relocate(species_l,.after=species)|> 
  separate(clim,into=c("var","quant"),remove=FALSE) |> 
  relocate(var,.after=species_l) |> 
  left_join(df.range,by=c("species","var")) |> 
  filter(!is.na(species_l)) |> 
  mutate(dif=case_when(grepl("opt",clim)~100*abs(clim_obs-clim_niche)/range,
                       grepl("high",clim)~100*(clim_obs-clim_niche)/range,
                       grepl("low",clim)~100*(clim_niche-clim_obs)/range)) |> 
  ggplot(aes(dif,var,color=var,shape=quant))+
  geom_point()+
  theme_minimal()+
  geom_vline(xintercept=0)+
  facet_wrap(~species)
df.tree.range|> 
  left_join(df.niche,by=c("species","clim")) |> 
  relocate(species_l,.after=species)|> 
  separate(clim,into=c("var","quant"),remove=FALSE) |> 
  relocate(var,.after=species_l) |> 
  left_join(df.range,by=c("species","var")) |> 
  filter(!is.na(species_l)) |> 
  mutate(dif=case_when(grepl("opt",clim)~100*abs(clim_obs-clim_niche)/range,
                       grepl("high",clim)~100*(clim_obs-clim_niche)/range,
                       grepl("low",clim)~100*(clim_niche-clim_obs)/range)) |> 
  ggplot(aes(dif,var,color=var,shape=quant))+
  geom_point()+
  theme_minimal()+
  geom_vline(xintercept=0)+
  facet_wrap(~species)

(df.tree.range.1 |> 
  left_join(df.niche,by=c("species","clim")) |> 
  relocate(species_l,.after=species)|> 
  separate(clim,into=c("var","quant"),remove=FALSE) |> 
  relocate(var,.after=species_l) |> 
  left_join(df.range,by=c("species","var")) |> 
  filter(!is.na(species_l)) |> 
  mutate(dif=case_when(grepl("opt",clim)~100*abs(clim_obs-clim_niche)/range,
                       grepl("high",clim)~100*(clim_obs-clim_niche)/range,
                       grepl("low",clim)~100*(clim_niche-clim_obs)/range)) |> 
  filter(dif<(-50)) |> 
  select(species) |> unique())$species->sp.deleted
sp.select=setdiff(unique(df.tree$species),sp.deleted)
```
~seuil 50% -> reste 10esp

sorbus aucupari, quercus ilex, carpinus betulus, quercus robur, quercus petaea, abies alba, quercus pubescens : ok
lauris nobilis, acer nigrum, fagus sylt, frangula alnus
pinus pinaster, cedrus atlantica, pinus pinea, pnius sylvestris, picea abies : jsp?

# Preliminary analysis of diameter
```{r}
library(ggpointdensity)
df.tree |> 
  filter(!species %in% sp.deleted) |> 
  filter(fecEstMu<10000000&diam<200) |> 
  ggplot(aes(diam,fecEstMu,color=species))+
  geom_pointdensity()+
  geom_hline(yintercept=10000000)+
  scale_y_log10()+scale_x_log10()


df.tree |> 
  filter(!species %in% sp.deleted) |> 
  filter(fecEstMu<10000000&diam<200) |> 
  mutate(fecCat=cut(fecEstMu,
                    breaks=10),
         diamCat=cut(diam,
                     breaks=10)) |> 
  ggplot(aes(diamCat,fecEstMu,color=species))+
  geom_boxplot(outlier.color = NA)+
  scale_y_log10()

for (sp in unique(df.tree$species)){
  print(
    df.tree |> 
      # filter(!species %in% sp.deleted) |> 
      filter(species==sp) |> 
      filter(fecEstMu<10000000&diam<200) |> 
      mutate(fecCat=cut(fecEstMu,
                        breaks=10),
             diamCat=cut(diam,
                         breaks=10)) |> 
      ggplot(aes(fecCat,diam,color=species))+
      geom_boxplot(outlier.color = NA)
  )
}
```
# Random forest
```{r}
library(randomForest)
df.rf=df.tree |> 
  mutate(across(c("fecEstMu","diam","mat","tmin","map"),
                scale))
randomForest(fecEstMu~species+diam+map+mat+tmin,data=df.rf)->rf_model
plot(rf_model,type="simple")
```
# Analysis of fec var with distance from opt
```{r}
for (sp in unique(df.tree$species)){
  print(
    df.tree |> 
      filter(species==sp) |> 
      filter(fecEstMu<10000000&diam<200) |> 
      group_by(species,plotID,map,mat,tmin,treeID) |> 
      summarise(fecEstMu=mean(fecEstMu),
                diam=mean(diam)) |>
      mutate(fecEstMuSt=fecEstMu/(diam^2)) |> 
      # select(treeID,plotID,year,species,diam,fecEstMuSt,map,mat,tmin) |> 
      pivot_longer(cols=c("map","mat","tmin"),names_to = "clim") |> 
      left_join(df.niche.short) |> 
      mutate(dist=(value-opt)/range) |> 
      ggplot(aes(dist,fecEstMuSt))+
      geom_point()+
      facet_wrap(~clim)
  )
}

```


# Explo IFN
```{r}
library(terra)
library("factoextra")
list.files("data/europe")->files
lapply(paste0("data/europe/",files),load,.GlobalEnv)


files.climate=list(mat="data/CHELSA/CHELSA_bio10_1.tif",
                   tmin="data/CHELSA/CHELSA_bio10_6.tif",
                   map="data/CHELSA/CHELSA_bio10_12.tif",
                   sgdd="data/CHELSA/CHELSA_gdd5_1981-2010_V.2.1.tif",
                   pet="data/CHELSA/CHELSA_pet_penman_mean_1981-2010_V.2.1.tif")
clim=rast(lapply(names(files.climate),
                 function(z){
                   r=rast(files.climate[[z]])
                   names(r)=z
                   return(r)
                   }
                   )
            )
plotData<- cbind(plotData,
                 extract(clim,
                         y=data.frame(x=plotData$lon,
                                      y=plotData$lat))[,-1]) |> 
  mutate(wai=(map-12*pet)/(12*pet))
rm(clim,files.climate,files)
# Perform PCA on the data
clim_pca <- prcomp(plotData[,c("sgdd","wai")] |> drop_na(),
                   scale. = TRUE) # graph = FALSE to suppress automatic visualization

# Plot the variables in the first PCA pane
fviz_pca_var(clim_pca, col.var = "contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, ggtheme = theme_classic())

# Make predictions of all observations on the first PCA axis
clim_pred <- cbind(plotData,
                   predict(clim_pca, newdata = plotData[,c("sgdd","wai")])) |> 
  rownames_to_column() |> 
  select(-rowname)

# Plot the observations on the first PCA axis
ggplot(data.frame(x = clim_pred[,"PC1"]), aes(x = x)) + 
  geom_density(fill = "#69b3a2", alpha = 0.5) + 
  xlab("First PCA Axis") + ylab("Density")

# format ISP
ISP_select<-as.data.frame(ISP)  |> 
  rownames_to_column(var="plot") |> 
  select(plot,matches(sp.select)) |> 
  pivot_longer(cols=-plot,
               names_to = "species",
               values_to = "ISP") |> 
  left_join(clim_pred)
BA_select<-as.data.frame(BA)|> 
  rownames_to_column(var="plot") |> 
  select(plot,matches(sp.select)) |> 
  pivot_longer(cols=-plot,
               names_to = "species",
               values_to = "BA") |> 
  left_join(clim_pred)
clim_range <- BA_select |> 
  filter(BA!=0) |> 
  group_by(species) |> 
  summarise(quant95=quantile(PC1,probs=0.95,na.rm=TRUE),
            quant525=quantile(PC1,probs=0.525,na.rm=TRUE),
            quant475=quantile(PC1,probs=0.475,na.rm=TRUE),
            quant05=quantile(PC1,probs=0.05,na.rm=TRUE))
ISP_select |> 
  left_join(clim_range) |> 
  mutate(margin=case_when(PC1<quant05~"low",
                          PC1>quant475&PC1<quant525~"opt",
                          PC1>quant95~"up",
                          TRUE~NA)) |> 
  filter(!is.na(margin)) |> 
  filter(ISP!=0) |>
  ggplot(aes(margin,ISP))+
  geom_boxplot()+ #outlier.colour = NA
  facet_wrap(~species,scales="free")+
  scale_y_log10()

```
# Improvement :
- weight by uncertainty
- sensitivity to quantiles delineation
- plot margins

## raw data of fecundity, sd, and basal area
```{r getdata}
load("data/europe/fecGmSd.rdata")
fecGmSd<-as.data.frame(summary)|> 
  rownames_to_column(var="plot") |> 
  select(plot,matches(sp.select)) |> 
  pivot_longer(cols=-plot,
               names_to = "species",
               values_to = "fecGmSd") |> 
  filter(species!="sorbusAucuparia") |> 
  left_join(clim_pred) 
load("data/europe/fecGmMu.rdata")
fecundity<-as.data.frame(summary)|> 
  rownames_to_column(var="plot") |> 
  select(plot,matches(sp.select)) |> 
  pivot_longer(cols=-plot,
               names_to = "species",
               values_to = "fecGmMu") |> 
  filter(species!="sorbusAucuparia") |> 
  left_join(fecGmSd) |> 
  left_join(BA_select) |> 
  left_join(clim_range)
rm(fecGmSd)
```

## weighted boxplot
Weighting obersvation do not drastically changes distribution.
```{r}
fecundity |> 
  mutate(fecGmCv=100*fecGmSd/fecGmMu,
         ISP=fecGmMu/BA,
         margin=case_when(PC1<quant05~"low",
                          PC1>quant475&PC1<quant525~"opt",
                          PC1>quant95~"up",
                          TRUE~NA)) |> 
  filter(!is.na(margin)) |> 
  filter(ISP!=0 | BA !=0) |> 
  filter(!is.nan(fecGmCv)) |> 
  # ggplot(aes(margin,ISP))+
  # geom_boxplot(outlier.colour = NA)+
  # facet_wrap(~species,scales="free")+
  # scale_y_log10()
  ggplot()+
  geom_density(aes(x=ISP,y=..density..,weight=fecGmCv,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_x_log10()+
  theme_minimal()

```

## plot margins per species

```{r}
#plot pc1
library(viridis)
fecundity |> 
  filter(species=="fagusSylvatic") |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=PC1,alpha=0.4))+
  xlim(-10,33)+ylim(37,70)+
  theme_minimal()+
  theme(axis.text = element_blank(),
        axis.title=element_blank())+
  scale_color_gradientn(colours = viridis(15))

# plot margin obs
fecundity |> 
  mutate(fecGmCv=100*fecGmSd/fecGmMu,
         ISP=fecGmMu/BA,
         margin=case_when(PC1<quant05~"low",
                          PC1>quant475&PC1<quant525~"opt",
                          PC1>quant95~"up",
                          TRUE~NA)) |> 
  filter(!is.na(margin)) |>
  filter(ISP!=0 | BA !=0) |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=margin,alpha=0.4))+
  xlim(-10,33)+ylim(37,70)+
  theme_minimal()+
  theme(axis.text = element_blank(),
        axis.title=element_blank())+
  facet_wrap(~species)
```

## Compute statistics of presence per specie
```{r}
fecundity |> 
  filter(BA!=0) |> 
  group_by(species) |> 
  summarise(n=n()) |> arrange(n)
```
## Bootstrap like
Idée : pour chaque espèce, sampling aléatoire d'un nombre n d'individus (ou d'un pourcentage du dataset), petit. Extraction des valeurs de ISP associés au quant975, 025 et 50 de PC1

```{r}
df.raw<-data.frame(species=NA,
                   BA=NA,
                   fecGmMu=NA,
                   fecGmSd=NA,
                   name=NA)

for (i in 1:500){
  print(i)
  try<-fecundity |> 
    select(plot,species,BA,fecGmMu,fecGmSd,lon,lat,PC1) |> 
    filter(BA!=0) |> 
    group_by(species) |> 
    sample_frac(0.1) |>
    mutate(quant95=abs(PC1-quantile(PC1,probs=0.95,na.rm=TRUE)[[1]]),
           quant5=abs(PC1-quantile(PC1,probs=0.5,na.rm=TRUE)[[1]]),
           quant05=abs(PC1-quantile(PC1,probs=0.05,na.rm=TRUE)[[1]])) |>
    pivot_longer(cols=c("quant95","quant5","quant05")) |> 
    ungroup() |> group_by(species,name) |> slice(which.min(value)) |> 
    select(species,BA,fecGmMu,fecGmSd,name)
  df.raw=rbind(df.raw,try)
}

df.raw |> 
  filter(!is.na(species)) |> 
  mutate(fecGmCv=100*fecGmSd/fecGmMu,
         ISP=fecGmMu/BA) |> 
  ggplot()+
  geom_density(aes(x=ISP,y=..density..,weight=fecGmCv,color=name))+
  facet_wrap(~species,scales="free")+
  scale_x_log10()+
  theme_minimal()

  
```
## sensitivity to quantiles
```{r}
prob.low=seq(0.01,0.47,0.01)
prob.high=seq(0.99,0.53,-0.01)
sensibility<-data.frame(species=unique(fecundity$species),
                        low.mid=NA,
                        low.high=NA,
                        high.mid=NA,
                        low.mid.sd=NA,
                        low.high.sd=NA,
                        high.mid.sd=NA) |> crossing(i=1:47)
for (i in 1:47){
  print(i)
  clim_i=BA_select |> 
    filter(!species%in%c("sorbusAucupari","sorbusAucuparia","cedrusAtlantic")) |> 
    filter(BA!=0) |> 
    group_by(species) |> 
    summarise(quant.high=quantile(PC1,probs=prob.high[i],na.rm=TRUE)[[1]],
              quant525=quantile(PC1,probs=0.525,na.rm=TRUE)[[1]],
              quant475=quantile(PC1,probs=0.475,na.rm=TRUE)[[1]],
              quant.low=quantile(PC1,probs=prob.low[i],na.rm=TRUE)[[1]])
  fecundity_i<-fecundity |> 
    select(plot,species,fecGmMu,fecGmSd,PC1,BA) |> 
    left_join(clim_i)|> 
    mutate(fecGmCv=100*fecGmSd/fecGmMu,
           ISP=fecGmMu/BA,
           margin=case_when(PC1<quant.low~"low",
                          PC1>quant475&PC1<quant525~"opt",
                          PC1>quant.high~"up",
                          TRUE~NA)) |> 
    filter(!is.na(margin)) |>
    filter(ISP!=0 | BA !=0) |>
    filter(!is.nan(fecGmCv))
  for (sp in unique(clim_i$species)){
    fecundity_i |> filter(species==sp)->test
    sensibility[sensibility$species==sp&sensibility$i==i,c("low.mid","low.high","high.mid")]=
      list(
        t.test(test[test$margin=="low","ISP"],test[test$margin=="opt","ISP"])$p.value,
        t.test(test[test$margin=="low","ISP"],test[test$margin=="up","ISP"])$p.value,
        t.test(test[test$margin=="up","ISP"],test[test$margin=="opt","ISP"])$p.value
      )
  }
  }
sensibility |> 
  pivot_longer(cols=c("low.mid","low.high","high.mid")) |> 
  ggplot(aes(i,value,color=name))+
    geom_line()+
    facet_wrap(~species)
```

