---
title: "Fecundity variability in species niche and distribution - outlines "
author: "Anne Baranger"
date: "2023-04-14"
output:
  bookdown::html_document2: 
    toc: TRUE 
  bookdown::pdf_document2: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE,message=FALSE)
library(targets)
lapply(c("stringr","ggplot2","tidyr","dplyr","rworldmap","modi","tibble","terra","factoextra","sf"),require,character.only=TRUE)
theme_set(theme_minimal())
tar_load(mastif.eu,store="target_data")
tar_load(mastif.am,store="target_data")

#tar_load(meanClimate_species,store="target_gbif")

tar_load(species_selection,store="target_nfi")

df.mastif.tree=rbind(mastif.am$df.tree,mastif.eu$df.tree) 
df.species=rbind(mastif.am$df.species.select,mastif.eu$df.species.select)
sp.select.eu=species_selection$sp.select.eu
sp.select.am=species_selection$sp.select.am


# small format of gbif
df.gbif.short <- species_selection$df.gbif |>
  separate(clim,into = c("clim","range")) |> 
  pivot_wider(names_from = range,
                values_from = clim_niche)

# tar_load(nfi.am_clim,store="target_nfi")
# tar_load(nfi.eu_clim,store="target_nfi")
tar_load(fecundity.eu_clim,store="target_nfi")
tar_load(fecundity.am_clim,store="target_nfi")
```

# Data description
species_selection 
 - df.gbif - niche by species, reformated for use with mastif
 - df.tree_w all tree (in df.tree) with computed weight
 - mastif.range - niche according to mastif plots with the use of weight
 - sp.select.am
 - sp.select.eu

df.nfi.eu/am -  plot of nfi, with size, climate

nfi.eu/am_clim
 - acp of climate variables
 - plots of nfi + etc + predictions of pca first axis

fec.eu/am
 - for each nfi plot, ba/ISP/fecmu/sd and weight of each plot -> scaled to the hectare, long format. all species represented on each plots even if absent

sp.margin.eu/am
 - for each species of the nfi, compute the quantile of pca first axi

fecundity.eu_clim
 - fecundity coupled with climate of plots, and condition of margin

-> fecundity datasets are long formats of nfi plots crossed by species list. It includes non-occurence of species.


# Mastif plots distribution
In this first section we use predictions performed on Mastif calibration plots. We filtered only plots where native european species were growing. In the following we present a map of the 1346 plots sampled in the different Mastif census. We also map plots specific to _Pinus sylvestris_.

```{r MastifPlots, fig.cap="Maps of Mastif calibration plots"}

# df.niche=read.csv("output/sp_gbif_climate.csv",sep=" ")
#plot initial dataset and selection
worldmap <- sf::st_as_sf(getMap(resolution = "high"))


# plots of selected species
df.mastif.tree |> 
  filter(species %in% df.species$species) |> 
  select(species,lon,lat) |> 
  distinct() |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=species))+
  xlim(range(df.species[,c("minlon","maxlon")])) + 
  ylim(range(df.species[,c("minlat","maxlat")]))+
  theme_minimal()+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "none")+
  labs(color="Species")
```


```{r MastifPlotsSp,fig.Cap="Mastif plots of Pinus sylvestris"}
df.mastif.tree |> 
  select(species,lon,lat) |>
  filter(species=="pinusSylvestr") |> 
  distinct() |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=species))+
  xlim(range(df.species[,c("minlon","maxlon")])) + 
  ylim(range(df.species[,c("minlat","maxlat")]))+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom")+
  labs(color="Species")

rm(worldmap)

```

# Comparing GBIF quantiles and Mastif distribution 

To perform our analysis, we cannot extrapolate Mastif to IFN plots ranging outside the gradient spanned by the mastif plots. Using GBIF and an R routine from GK and JB, we computed 5th and 95th quantiles over 3 climate variables : tmin, map, and mat.
We computed on Mastif database the same quantile, using the plot distribution. However, because data were not collected homogenously, we weighted each observations by the coefficient of variation computed from fecundity predictions and stand error.




Figure  \@ref(fig:ClimateDifFig) presents the difference percentage between GBIF quantiles and wieghted quantiles of Mastif.

```{r ClimateDifFig, fig.cap="Climate difference between GBIF and Mastif quantiles"}
# select ranges computed from mastif plots (with weight)
species_selection$mastif.range |> 
  # merge with gbif range
  left_join(species_selection$df.gbif,by=c("species","clim")) |> 
  filter(!is.na(species_l)) |>  # filter for absent species 
  relocate(species_l,.after=species) |> 
  mutate(dif=100*(clim_obs-clim_niche)/clim_niche) |> 
  ggplot(aes(dif,clim,color=species))+
  geom_point()+
  theme(legend.position = "none")


```


On the following histograms we overlayed the weighted observations of Mastif and the GBIF range (5th and 95th quantiles, and median) for species that were not selected by our algorithm. This enables to glimpse how different are Mastif plots distribution compared to GBIF.


```{r PlotsnSp,fig.cap="Numbers of plots per species"}
df.mastif.tree |> 
  filter(!duplicated(treeID)) |> 
  group_by(species) |> 
  mutate(n_ind=n()) |> ungroup() |> 
  select(species,n_ind,plotID) |> 
  unique() |> 
  group_by(species,n_ind) |> 
  summarise(n=n())


```

Test difference in quantiles and optima, between GBIF and MASTIF :
```{r}
#confront mastif climate quantiles and gbif
species_selection$mastif.range |> 
  separate(clim,into=c("var","quant"),remove=FALSE) |> 
  left_join(df.gbif.short,by=c("species","var"="clim")) |> 
  relocate(species_l,.after=species)|> 
  filter(!is.na(species_l)) |> 
  pivot_longer(cols=c("low","opt","high"),
               names_to = "quant_niche",
               values_to = "clim_niche") |> 
  filter(quant==quant_niche) |>  select(-quant_niche) |> 
  mutate(dif=case_when(grepl("opt",clim)~100*abs(clim_obs-clim_niche)/range,
                       grepl("high",clim)~100*(clim_obs-clim_niche)/range,
                       grepl("low",clim)~100*(clim_niche-clim_obs)/range)) |> 
  ggplot(aes(dif,var,color=var,shape=quant))+
  geom_point()+
  theme_minimal()+
  geom_vline(xintercept=0)+
  facet_wrap(~species)
```

We set a threshold of -50% of difference. We ended up selecting the following species for further analysis:
```{r SpSelect}
sp.select.am
sp.select.eu
```

## diagnosis of deleted species
```{r}
sp.deleted=unique(df.species$species)[!unique(df.species$species)%in% c(sp.select.am,sp.select.eu)] 
for (sp in sp.deleted){
  plot<-species_selection$df.tree_w |>
    select(plotID,species,clim,clim_val,cv_plot,low,high) |> 
    left_join(df.gbif.short,by=c("species","clim")) |> 
    filter(species==sp) |> 
    ggplot()+
    geom_histogram(aes(x=clim_val,y=..density..,weight=1/cv_plot))+
    geom_point(aes(x=opt,y=0),color="darkred",size=1)+
    geom_segment(aes(x=low.y,xend=high.y,y=-0,yend=-0),color="darkred",size=1)+
    geom_vline(aes(xintercept = low.x))+
    geom_vline(aes(xintercept = high.x))+
    facet_wrap(~clim,
               # nrow = 3,
               scales = "free")+
    labs(title=sp)
  print(plot)
  rm(plot)
}
rm(sp.deleted)
```

## Species selection with order statistics

```{r orderstats}
#format data with order stats
df.tree=rbind(mastif.am$df.tree,mastif.eu$df.tree) |>
    # replace 0 estimations by NA and compute, rescale mat and tmin
    mutate(fecEstSe=na_if(fecEstSe,0),
           fecEstMu=na_if(fecEstMu,0),
           across(c("mat","tmin"),
                  ~.x/10)
    ) |> 
    filter(!is.na(fecEstMu)) |> filter(!is.na(fecEstSe)) |> 
    # compute numbers of observations per plot
    group_by(plotID,species,mat,map,tmin) |> 
    summarise(n_plot=n(),
              fecEstMu_plot=mean(fecEstMu,na.rm=TRUE),
              se_plot=sqrt(sum(fecEstSe^2)/n_plot),
              cv_plot=se_plot/fecEstMu_plot) |> 
    # Compute mean of se per plot
    ungroup() |> 
    pivot_longer(cols=c("mat","tmin","map"),
                 names_to = "clim",
                 values_to = "clim_val")  |> 
    filter(!is.na(clim_val)) |> 
    group_by(species,clim) |>
    mutate(opt=weighted.quantile(clim_val,1/cv_plot,prob=0.5)[1],
           low=weighted.quantile(clim_val,1/cv_plot,prob=0.025)[1],#,probs=0.1)[1],
           high=weighted.quantile(clim_val,1/cv_plot,prob=0.975)[1],
           high.rank=quantile(clim_val, probs = 1- min(10, n())/n(), na.rm = TRUE),
           low.rank=quantile(clim_val, probs = min(10, n())/n(), na.rm = TRUE)) |> 
    ungroup()
# format data for matching
df.tree |>
    # format to match df.niche 
    select(species,clim,low,low.rank,high,high.rank) |> unique() |> 
    pivot_longer(cols=c("low","high"),names_to = "quant",values_to = "quant_val") |> 
    pivot_longer(cols=c("low.rank","high.rank"),names_to = "rank",values_to = "rank_val") |> 
    filter((quant=="low"&rank=="low.rank")|
             (quant=="high"&rank=="high.rank"))
    pivot_wider(names_from = "clim",
                values_from = c("low.rank","high.rank","high","low"),
                names_sep = ".") |> 
    pivot_longer(cols=-species,
                 names_to = "clim",
                 values_to = "clim_obs") |> 
    mutate(clim=paste0(str_split_i(clim,"\\.",2),".",str_split_i(clim,"\\.",1))) -> mastif.range
#compute climate range
df.range<-species_selection$df.gbif |> 
    select(species,clim,clim_niche) |> 
    filter(grepl("range",clim)) |> 
    separate(clim,into=c("var","drop")) |> 
    select(-drop) |> 
    rename(range="clim_niche")

# create distribution stat for mastif
mastif.range=df.tree |> 
  select(species,clim,low,low.rank,high,high.rank) |> unique() |> 
  pivot_longer(cols=c("low","high"),names_to = "quant",values_to = "quant_val") |> 
  pivot_longer(cols=c("low.rank","high.rank"),names_to = "rank",values_to = "rank_val") |> 
  filter((quant=="low"&rank=="low.rank")|
         (quant=="high"&rank=="high.rank")) |> 
  mutate(clim=paste0(clim,".",quant)) |> 
  select(-quant,-rank) |> 
  pivot_longer(cols = c("quant_val","rank_val"),
               names_to = "stat",
               values_to = "clim_obs" ) |> 
  separate(stat,into = c("stat","val")) |> select(-val) 

#,plot dif between quant and order
mastif.range |> 
  pivot_wider(names_from = stat,
              values_from = clim_obs) |>
  separate(clim,into=c("clim","edge")) |> 
  ggplot(aes(quant,rank,colour=species,shape=edge))+
  geom_point()+
  theme(legend.position = "none")+
  geom_abline(slope=1)+
  facet_wrap(~clim,scales="free")

# species selection
mastif.range|> 
    left_join(species_selection$df.gbif,by=c("species","clim")) |> 
    relocate(species_l,.after=species)|> 
    separate(clim,into=c("var","quant"),remove=FALSE) |> 
    relocate(var,.after=species_l) |> 
  # merge climate range
    left_join(df.range,by=c("species","var")) |> 
    filter(!is.na(species_l)) |> 
  # filter(stat!="rank") |> 
    mutate(dif=case_when(grepl("high",clim)~100*(clim_obs-clim_niche)/range,
                         grepl("low",clim)~100*(clim_niche-clim_obs)/range)) |>
    mutate(dif.cat=cut(dif,breaks=c(0,-10,-20,-30,-40,-50,-60,-70,-80,-90,-100))) |> 
  group_by(species,stat,dif.cat) |>
    summarise(n=n()) |>
  pivot_wider(names_from=stat,
              values_from = n)
  ggplot(aes(dif.cat,n,colour=species))+
    theme(legend.position = "none")+
  geom_point()+
  facet_wrap(~stat)

```




# Preliminary analysis of diameter

We explored the link between diameter and fecundity. We assumed fecundity increased with higher diameter, and decided to standardize the fecundity estimates by the basal area of the producing tree.
```{r}

# plot of diameter/fecundity coloured by species
# library(ggpointdensity)
# df.tree |> 
#   filter(!species %in% sp.deleted) |> 
#   filter(fecEstMu<10000000&diam<200) |> 
#   ggplot(aes(diam,fecEstMu,color=species))+
#   geom_pointdensity()+
#   geom_hline(yintercept=10000000)+
#   scale_y_log10()+scale_x_log10()

# boxplot of fecundity by diameter categories, coloured by species


# df.tree |> 
#   filter(!species %in% sp.deleted) |> 
#   filter(fecEstMu<10000000&diam<200) |> 
#   mutate(fecCat=cut(fecEstMu,
#                     breaks=10),
#          diamCat=cut(diam,
#                      breaks=10)) |> 
#   ggplot(aes(diamCat,fecEstMu,color=species))+
#   geom_boxplot(outlier.color = NA)+
#   scale_y_log10()

for (sp in sample(unique(df.mastif.tree$species),10)){
  print(
    df.mastif.tree |> 
      # filter(!species %in% sp.deleted) |> 
      filter(species==sp) |> 
      filter(fecEstMu<10000000&diam<200) |> 
      mutate(fecCat=cut(fecEstMu,
                        breaks=10),
             diamCat=cut(diam,
                         breaks=10)) |> 
      ggplot(aes(fecCat,diam,color=species))+
      geom_boxplot(outlier.color = NA)
  )
}
```


<!-- # Random forest -->
```{r eval=FALSE, include=FALSE}
library(randomForest)
df.rf=df.mastif.tree |> 
  mutate(across(c("fecEstMu","diam","mat","tmin","map"),
                scale))
randomForest(fecEstMu~species+diam+map+mat+tmin,data=df.rf)->rf_model
plot(rf_model,type="simple")
```


<!-- # Analysis of fec var with distance from opt -->
```{r eval=FALSE, include=FALSE}
for (sp in unique(df.mastif.tree$species)){
  print(
    df.mastif.tree |> 
      filter(species==sp) |> 
      filter(fecEstMu<10000000&diam<200) |> 
      group_by(species,plotID,map,mat,tmin,treeID) |> 
      summarise(fecEstMu=mean(fecEstMu),
                diam=mean(diam)) |>
      mutate(fecEstMuSt=fecEstMu/(diam^2)) |> 
      # select(treeID,plotID,year,species,diam,fecEstMuSt,map,mat,tmin) |> 
      pivot_longer(cols=c("map","mat","tmin"),names_to = "clim") |> 
      left_join(df.niche.short) |> 
      mutate(dist=(value-opt)/range) |> 
      ggplot(aes(dist,fecEstMuSt))+
      geom_point()+
      facet_wrap(~clim)
  )
}

```


# Analysis of fecundity variation in species niche margins and optima, using European FNI

## Climate ACP and boxplot of fecundity for few species

```{r}
# Plot the variables in the first PCA pane for America
tar_load(c(nfi.eu_clim,nfi.am_clim),store="target_nfi")
fviz_pca_var(nfi.am_clim$clim_pca, col.var = "contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, ggtheme = theme_classic())

# Plot the variables in the first PCA pane for Europe
fviz_pca_var(nfi.eu_clim$clim_pca, col.var = "contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, ggtheme = theme_classic())
rm(nfi.eu_clim,nfi.am_clim)

# first plot with no weight
for (dataset in list(fecundity.am_clim,fecundity.eu_clim)){
  print(
  dataset |> 
    filter(!is.na(margin)) |> 
    filter(ISP!=0) |>
    # filter(species %in% sample(sp.select.am,10)) |> 
    ggplot(aes(margin,ISP))+
    geom_boxplot()+ #outlier.colour = NA
    facet_wrap(~species,scales="free")+
    scale_y_log10()
  )
}

```
## Margins plots

```{r}
#plot pc1
library(viridis)
worldmap <- sf::st_as_sf(getMap(resolution = "high"))

fecundity.eu_clim |> 
  filter(species=="fagusSylvatic") |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=PC1,alpha=0.4))+
  xlim(-10,33)+ylim(37,70)+
  theme_minimal()+
  theme(axis.text = element_blank(),
        axis.title=element_blank())+
  scale_color_gradientn(colours = viridis(15))

# plot margin obs
fecundity.eu_clim |> 
  filter(!is.na(margin)) |>
  filter(ISP!=0 | BA !=0) |>
  filter(species %in% c("abiesAlba","fagusSylvatic")) |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=margin,alpha=0.4))+
  xlim(-10,33)+ylim(37,70)+
  theme_minimal()+
  theme(axis.text = element_blank(),
        axis.title=element_blank())+
  facet_wrap(~species)

# plot all low margin
fecundity.eu_clim |> 
  filter(margin=="low") |> 
  filter(ISP!=0 | BA !=0) |>
  # filter(species %in% c("abiesAlba","fagusSylvatic")) |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=species,alpha=0.4))+
  xlim(-10,33)+ylim(37,70)+
  theme_minimal()+
  theme(axis.text = element_blank(),
        axis.title=element_blank())

```

## Climatic space of north america
```{r}
tar_load(df.nfi.am,store="target_nfi")
clim_list=list(mat="data/CHELSA/CHELSA_bio10_1.tif",
               tmin="data/CHELSA/CHELSA_bio10_6.tif",
               map="data/CHELSA/CHELSA_bio10_12.tif",
               sgdd="data/CHELSA/CHELSA_gdd5_1981-2010_V.2.1.tif",
               pet="data/CHELSA/CHELSA_pet_penman_mean_1981-2010_V.2.1.tif")
clim_list_sup=paste0("data/CHELSA/sup/",list.files("data/CHELSA/sup/"))
names(clim_list_sup)=substr(clim_list_sup,24,24+7)
clim_list=c(clim_list,clim_list_sup)
# clim=rast(lapply(names(clim_list),
#                  function(z){
#                    r=terra::rast(clim_list[[z]])
#                    names(r)=z
#                    return(r)
#                  }
#                  )
#           )
# 
for (z in names(clim_list_sup)){
   print(z)
   df.nfi.am <- cbind(df.nfi.am,
                      clim=extract(rast(clim_list[[z]]),
                              y=data.frame(x=df.nfi.am$lon,
                                           y=df.nfi.am$lat))[,-1]) |> 
     rename({{z}}:=clim)
  
}
df.acp=fecundity.am_clim |> filter(species=="fagusGrandifo") |> filter(BA!=0) |> select("sgdd","wai") |> drop_na()
clim_pca<-prcomp(df.acp,
                 scale. = TRUE)
clim_pca<-prcomp(df.nfi.am[,12:29] |> drop_na(),
                 scale. = TRUE) 
clim_pca<-prcomp(df.nfi.am[,c("vpd_max_","map","mat","sgdd","bio15_19","bio4_198","cmi_max_","bio14_19")] |> drop_na(),
                 scale. = TRUE)
clim_pca<-prcomp(df.nfi.am[species=="fagusGrandifo",c("sgdd","wai")] |> drop_na(),
                 scale. = TRUE)
fviz_pca_var(clim_pca, col.var = "contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, ggtheme = theme_classic())
fviz_pca_var(clim_pca, col.var = "contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), axes=c(2,3),
             repel = TRUE, ggtheme = theme_classic())
fviz_pca_ind(clim_pca, geom.ind = "point")
rm(df.nfi.am)


cbind(df.nfi.am,
      predict(clim_pca, newdata = df.nfi.am[,c("sgdd","wai")])) |>
  rownames_to_column() |>
  select(-rowname) |> 
  mutate(margin=case_when(PC1<quantile(PC1,probs=0.05,na.rm=TRUE)[[1]]~"coldhumid",
                          PC1>quantile(PC1,probs=0.95,na.rm=TRUE)[[1]]~"hotdry",
                          (PC1<quantile(PC1,probs=0.5,na.rm=TRUE)[[1]]&
                            PC1>quantile(PC1,probs=0.4,na.rm=TRUE)[[1]]&
                            PC2<quantile(PC1,probs=0.5,na.rm=TRUE)[[1]]&
                            PC2>quantile(PC1,probs=0.4,na.rm=TRUE)[[1]])~"core",
                          PC2>quantile(PC2,probs=0.95,na.rm=TRUE)[[1]]~"hothumid")) |> 
   ggplot(aes(PC1,PC2,color=margin))+geom_point()
  # group_by(margin) |> summarise(n=n())

fecundity.am_clim |> 
  filter(BA!=0) |> 
  group_by(species) |> 
  mutate(margin=case_when(PC1<quantile(PC1,probs=0.05,na.rm=TRUE)[[1]]~"coldhumid",
                          PC1>quantile(PC1,probs=0.95,na.rm=TRUE)[[1]]~"hotdry",
                          (PC1<quantile(PC1,probs=0.5,na.rm=TRUE)[[1]]&
                            PC1>quantile(PC1,probs=0.4,na.rm=TRUE)[[1]]&
                            PC2<quantile(PC1,probs=0.5,na.rm=TRUE)[[1]]&
                            PC2>quantile(PC1,probs=0.4,na.rm=TRUE)[[1]])~"core",
                          PC2>quantile(PC2,probs=0.95,na.rm=TRUE)[[1]]~"hothumid")) |> 
  ungroup() |> 
  filter(!is.na(margin)) |> 
  filter(ISP!=0) |>
  ggplot()+
  # geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
  geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_x_log10()+
  theme_minimal()
```


## Weighted boxplot

### Europe
Weighting obersvation do not drastically changes distribution.
```{r}
# graph densité fécondité corrigée a chaque marge 
fecundity.eu_clim |> 
  filter(!is.na(margin)) |> 
  filter(ISP!=0) |>
  ggplot()+
  # geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
  geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_x_log10()+
  theme_minimal()

# graph densité fécondité non corrigée a chaque marge 
fecundity.eu_clim |> 
  filter(!is.na(margin)) |> 
  filter(ISP!=0) |>
  ggplot()+
  geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
  # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_y_log10()+
  theme_minimal()

fecundity.eu_clim |> 
  filter(!is.na(margin)) |> 
  filter(ISP!=0) |>
  ggplot()+
  geom_boxplot(aes(margin,cv_plot,color=margin),outlier.colour = NA)+
  # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_y_log10()+
  theme_minimal()



```
#### t-test

```{r compmed}
sp.dataset=as.data.table(fecundity.eu_clim |> filter(BA!=0))
for (sp in unique(fecundity.eu_clim$species)){
  print(sp)
  print(t.test(sp.dataset[species==sp&margin=="up",ISP],sp.dataset[species==sp&margin=="opt",ISP])
  )
  print(t.test(sp.dataset[species==sp&margin=="low",ISP],sp.dataset[species==sp&margin=="opt",ISP])
  )
  print(
    fecundity.eu_clim |>
      filter(species==sp) |> 
      filter(!is.na(margin)) |> 
      filter(ISP!=0) |>
      ggplot()+
      # geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
      geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
      geom_vline(xintercept = mean(sp.dataset[species==sp&margin=="low",ISP]),color="red")+
      geom_vline(xintercept = mean(sp.dataset[species==sp&margin=="up",ISP]),color="blue")+
      geom_vline(xintercept = mean(sp.dataset[species==sp&margin=="opt",ISP]),color="green")+
      scale_x_log10()+
      theme_minimal()+
      ggtitle(sp)
)
}

```

#### KS test :

```{r}
for (sp in unique(fecundity.eu_clim$species)){
  print(sp)
  sp.density=fecundity.eu_clim |> 
      filter(species==sp) |> 
      filter(!is.na(margin)) |> 
      filter(ISP!=0) |> 
    as.data.table()
  print(
     sp.density|>
      ggplot()+
      # geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
      geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
      facet_wrap(~species,scales="free")+
      scale_x_log10()+
      theme_minimal())
  optlow=ks.test(sp.density[margin=="low",ISP],sp.density[margin=="opt",ISP])
  lowup=ks.test(sp.density[margin=="low",ISP],sp.density[margin=="up",ISP])
  optup=ks.test(sp.density[margin=="up",ISP],sp.density[margin=="opt",ISP])
  print(
    paste0(
      c("optlow D : ","optlow pval : "),
      c(D=optlow$statistic,pval=optlow$p.value)
    )
    
  )
  print(
    paste0(
      c("lowup D : ","lowup pval : "),
      c(D=lowup$statistic,pval=lowup$p.value)
    )
    
  )
    print(
    paste0(
      c("optup D : ","optup pval : "),
      c(D=optup$statistic,pval=optup$p.value)
    )
    
  )
}
```



### Northern America 
```{r}
# graph densité fécondité corrigée a chaque marge 
fecundity.am_clim |> 
  filter(!is.na(margin)) |> 
  filter(ISP!=0) |>
  ggplot()+
  # geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
  geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_x_log10()+
  theme_minimal()

# graph densité fécondité non corrigée a chaque marge 
fecundity.am_clim |> 
  filter(!is.na(margin)) |> 
  filter(ISP!=0) |>
  ggplot()+
  geom_boxplot(aes(margin,ISP,color=margin),outlier.colour = NA)+
  # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_y_log10()+
  theme_minimal()

```


## ISP/margins interspecific

```{r eval=FALSE, include=FALSE}
min((fec_trans |> 
  group_by(species) |> 
  summarise(n=n()) |> 
  arrange(n))$n)->min.n

fec_trans |> 
  group_by(species) |> 
  sample_n(min.n) |> 
  ggplot()+
  geom_boxplot(aes(margin,ISP,color=margin),outlier.colour = NA)+
  scale_y_log10()+
  theme_minimal()
```


## Test using a continous distance - Mahanolinis distance

```{r}
sp="abiesAlba"
for(sp in unique(fecundity.eu_clim$species)){
  fec.sp=fecundity.eu_clim |> 
    filter(species==sp) |> 
    filter(!is.na(BA)) |> 
    mutate(across(c("mat","tmin","map","sgdd","pet","wai"),
                  scale))

  # geographic centroid and distance
  points_sp<-st_union(st_as_sf(fec.sp[,c("lon","lat")],coords=c("lon","lat")))
  centroid_sp<-st_centroid(st_convex_hull(points_sp))
  # centroid_sp2<-st_centroid(points_sp)
  fec.sp=cbind(fec.sp,
               dist_geo=st_distance(st_as_sf(fec.sp[,c("lon","lat")],
                                             coords=c("lon","lat")),
                                    centroid_sp))
  
  print(
    fec.sp |> 
      ggplot(aes(log(ISP),dist_geo))+
      geom_point(alpha=0.4,stroke=NA)+
      geom_smooth(method="lm")+
      ggtitle(label = paste0("geo_",sp))
  )
  
  
  # estimate climate centroid and distance
  fec.sp<-fec.sp[,c("lon","lat","margin","dist_geo","BA","ISP","mat","wai","tmin","sgdd")] |> 
    filter(!is.na(mat)&!is.na(wai)&!is.na(tmin)&!is.na(sgdd))
  sp.elp=covMcd(fec.sp[,c("mat","wai","tmin","sgdd")],alpha=0.7)
  sp.center=c(as.matrix(sp.elp$center))
  
  fec.sp=cbind(fec.sp,
               dist_maha=mahalanobis(fec.sp[,c("mat","wai","tmin","sgdd")],
                                     sp.center,
                                     cov(fec.sp[,c("mat","wai","tmin","sgdd")] |> drop_na())))
  
  
  
  print(
    fec.sp |> 
      ggplot(aes(log(ISP),dist_maha))+
      geom_point(alpha=0.4,stroke=NA)+
      geom_smooth(method="lm")+
      ggtitle(label = paste0("clim_",sp))
  )
    
  print(
    fec.sp |>
      ggplot(aes(margin,dist_maha))+
      geom_boxplot()
  )

}
# fec.sp |> 
#   ggplot(aes(dist_geo,dist_maha))+geom_point()+geom_smooth()
# fec.sp |> 
#   ggplot(aes(margin,dist_maha))+geom_boxplot()
# fec.sp |> 
#   ggplot(aes(lat,log(BA)))+geom_point()+geom_smooth(method="lm")

```


# Fecundity variation in geographical space

First delineate geographical margins :

##  delineate on latitude criteria
```{r geomargin}
fecundity.eu_clim |> 
  filter(BA!=0) |> 
  group_by(species) |> 
  mutate(lat95=quantile(lat,probs=0.95,na.rm=TRUE)[[1]],
            lat525=quantile(lat,probs=0.525,na.rm=TRUE)[[1]],
            lat475=quantile(lat,probs=0.475,na.rm=TRUE)[[1]],
            lat05=quantile(lat,probs=0.05,na.rm=TRUE)[[1]],
         margin=case_when(lat<lat05~"low",
                          lat>lat475&lat<lat525~"opt",
                          lat>lat95~"up",
                          TRUE~NA)) |> 
  filter(!is.na(margin)) |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=species,alpha=0.4))+
  xlim(-10,33)+ylim(37,70)+
  theme_minimal()+
  theme(axis.text = element_blank(),
        axis.title=element_blank())+
  facet_wrap(~margin)


```
## delineate using distance to core population


```{r geomargin}
sp.dataset=as.data.table(fecundity.eu_clim |> filter(BA!=0))
sp.dataset<-sp.dataset[species==sp,]
sp.dataset |> 
  ggplot() +
  geom_point(aes(x=lon,y=lat,color=margin)) +
geom_sf(data=worldmap,fill=NA)+
  xlim(-10,33)+ylim(37,70)+
  theme_minimal()+
  theme(axis.text = element_blank(),
        axis.title=element_blank())  

# determine centroid
nmin=sp.dataset[BA != 0,.(n = .N),by = species][n == sort(n)[2],n]
for (sp in unique(sp.dataset$species)){  #c("sorbusAucupari","fagusSylvatic","abiesAlba","piceaAbies")
  print(sp)
  
  # compute centroid of distribution
  sp.geodataset <- st_as_sf(sp.dataset[species==sp,], coords = c("lon", "lat"), crs = 4326)
  sp.centroid <- st_centroid(st_union(sp.geodataset$geometry),ratio=1)
  # plot(sp.geodataset$geometry)
  # plot(sp.centroid,add=TRUE,col="red")
  # n=max(round(dim(sp.dataset[species==sp,])[1]/1000),9)
 
  # first select points far from centroid -> i.e 15% the furthest 
  df.kmeans=cbind(sp.dataset[species==sp,],
      dist=as.numeric(st_distance(sp.geodataset,sp.centroid))) |>
    sample_n(nmin,replace=TRUE)
    # filter(dist>quantile(dist,probs=0.85)[[1]])
    # arrange(desc(dist)) |>
    # slice(1:(max(500,n()/10)))
  
  # print distribution and overlay points selected for margin
  print(
    ggplot()+
      geom_point(data=sp.dataset[species==sp,],aes(x=lon,y=lat))+
      geom_point(data=df.kmeans,aes(x=lon,y=lat),color="red")+
      geom_sf(data=worldmap,fill=NA)+
      xlim(-10,33)+ylim(37,70) # europe
      # xlim(-125,-75)+ylim(24,70) # northamerica
    )
  
  # dbscan
  dbscan::kNNdistplot(df.kmeans[,c("lon","lat")], k =  7)
  eps=quantile(dbscan::kNNdist(df.kmeans[,c("lon","lat")], k =  7),probs=0.85)[[1]]
  print(eps)
  fpc::dbscan(df.kmeans[,c("lon","lat")],eps=eps,MinPts =7)->df.dbscan

  # print dbscan clusters
  print(
    cbind(df.kmeans,
          cluster=df.dbscan$cluster) |>
      filter(cluster!=0) |> 
      ggplot()+
      geom_sf(data=worldmap,fill=NA,alpha=0.4)+
      geom_point(aes(x=lon,y=lat,colour=as.factor(cluster)))+
      xlim(-10,33)+ylim(37,70) # europe
      # xlim(-125,-75)+ylim(24,70) # northamerica
  )
  
  
  # boxplots of fecundity in clusters with enough individuals
  print(
    cbind(df.kmeans,
          cluster=as.character(df.dbscan$cluster))|> 
      filter(cluster !=0) |> 
      group_by(cluster) |> 
      mutate(n.cluster=n(),
             mean.lat=mean(lat),
             mean.dist=mean(dist)) |> 
      ungroup() |>
      filter(n.cluster>40) |> 
      mutate(order.min=dense_rank(mean.dist)) |> 
      filter(order.min==1 | order.min > max(order.min)-3) |> 
      # filter(cluster=="core"|
      #          mean.lat==max(mean.lat)|
      #          mean.lat==min(mean.lat)) |> 
      mutate(cluster=forcats::fct_reorder(as.factor(cluster),mean.dist)) |> 
      ggplot(aes(cluster,ISP,fill=mean.lat))+
      geom_boxplot()+
      scale_y_log10()+
      ggtitle(label=sp)

  )
  }
  # dbscan::kNNdistplot(df.kmeans[,c("lon","lat")], k =  7)
  # 
  # # boxplots of fecundity in clusters with enough individuals
  # print(
  #   bind_rows(
  #     # core population
  #     cbind(sp.dataset[species==sp,],
  #           dist=as.numeric(st_distance(sp.geodataset,sp.centroid))) |> 
  #       arrange(dist) |>
  #       slice(1:(max(300,n()/15))) |> 
  #       mutate(cluster="core"),
  #     # clusters
  #     cbind(df.kmeans,
  #       cluster=as.character(df.dbscan$cluster))
  #   ) |> 
  #     filter(cluster !=0) |> 
  #     group_by(cluster) |> 
  #     mutate(n.cluster=n(),
  #            mean.lat=mean(lat),
  #            mean.dist=mean(dist)) |> 
  #     ungroup() |>
  #     filter(n.cluster>40) |> 
  #     mutate(order.min=dense_rank(mean.dist)) |> 
  #     filter(order.min==1 | order.min > max(order.min)-3) |> 
  #     # filter(cluster=="core"|
  #     #          mean.lat==max(mean.lat)|
  #     #          mean.lat==min(mean.lat)) |> 
  #     mutate(cluster=forcats::fct_reorder(as.factor(cluster),mean.dist)) |> 
  #     ggplot(aes(cluster,ISP,fill=mean.lat))+
  #     geom_boxplot()+
  #     scale_y_log10()
  # 
  # )

#   # code with k-means algo
#   # kmeans
#   kmeans(df.kmeans[,c("lon","lat")],4)->sum.kmeans
#   ## map clusters
#   print(cbind(df.kmeans,
#         cluster=sum.kmeans$cluster) |> 
#     ggplot(aes(x=lon,y=lat,colour=as.factor(cluster)))+
#     geom_point())
#   ## draw boxplots
#   print(cbind(df.kmeans,
#         cluster=sum.kmeans$cluster) |> 
#     group_by(cluster) |> 
#     mutate(m.dist=mean(dist)) |> 
#     ungroup() |> 
#     mutate(order.min=dense_rank(m.dist)) |> 
#     filter(order.min==1 | order.min > max(order.min)-3) |> 
#     mutate(cluster=forcats::fct_reorder(as.factor(cluster),m.dist)) |> 
#     ggplot(aes(as.factor(cluster),ISP,fill=m.dist))+
#     geom_boxplot()+
#     scale_y_log10())




```

## regression ISP/distance
```{r}
sp.dataset=as.data.table(fecundity.eu_clim |> filter(BA!=0))


# determine centroid
sp.dist=sp.dataset |> 
  st_as_sf(coords = c("lon", "lat"), crs = 4326) |> 
  group_by(species) |> 
  mutate(centroid=st_centroid(st_union(geometry))) |> 
  mutate(lat=st_coordinates(geometry)[,2],
         x_centroid=st_coordinates(centroid)[,1],
         y_centroid=st_coordinates(centroid)[,2]) |> 
  ungroup() |> 
  mutate(dist=as.numeric(st_distance(geometry,centroid,by_element = TRUE)))

sp.dist |> 
  group_by(species) |> 
  mutate(ISP=scale(ISP),
         dist=scale(dist)) |> 
  mutate(dist=cut(dist,
                  breaks = 10)) |> 
  ggplot(aes(dist,ISP))+
  geom_boxplot()+
  scale_y_log10()+
  facet_wrap(~species,scales="free")


# lat/dist plots
sp.dist |> 
  group_by(species) |> 
  mutate(scale=ISP) |>  
  mutate(dist=cut(dist,breaks=10),
         lat=cut(lat,breaks=10)) |> ungroup() |>
  group_by(species,dist,lat) |> 
  summarise(mean.fec=mean(ISP)) |> 
  ggplot(aes(x=dist,y=lat))+
  geom_tile(aes(fill=log(mean.fec)))+
  scale_fill_distiller(palette="YlGnBu",direction=1)+
  facet_wrap(~species,scales="free")



# regression 
fecdist.lm=lm(ISP~species+dist+lat+dist:species,data=sp.dist)
summary(fecdist.lm)
anova(fecdist.lm)


fecdist.rq <- rq(ISP~species+dist+lat+dist:species,data=sp.dist, tau = 0.95)
summary(fecdist.rq)
```

# Fecundity variability in negative safety margins

```{r sfm}
psi_min=rast(fread("../../Era5-land/output/psihorday_real.csv")[,c("x","y","psi")],crs="epsg:4326")
frost.index.winter=min(rast("../../Era5-land/data/CHELSA/CHELSA_EUR11_tasmin_month_min_19802005.nc"),na.rm=FALSE)
species.traits=readRDS("../../Era5-land/target_safetymargin/objects//df.traits") |> 
  mutate(sp8=paste0(substr(tolower(genus),1,8),substr(str_to_title(species),1,8))) 
mod.output=readRDS("../../Era5-land/target_analysis/objects/df.mod.select") |> 
  separate(species.binomial, into=c("genus","sp")) |> 
  mutate(species=paste0(substr(tolower(genus),1,8),substr(str_to_title(sp),1,8)),
         t_hsm=na_if(t_hsm,0),
         t_fsm=na_if(t_fsm,0)) 
  
fec.eu_sfm=cbind(fecundity.eu_clim,
                 psi=extract(psi_min,data.frame(x=fecundity.eu_clim$lon,
                                      y=fecundity.eu_clim$lat))[[2]],
                 frost=extract(frost.index.winter,data.frame(x=fecundity.eu_clim$lon,
                                      y=fecundity.eu_clim$lat))[[2]]) |> 
  left_join(species.traits[,c("sp8","px.mu","lt50.mean")],by=c("species"="sp8")) |> 
  left_join(mod.output[,c("t_fsm","t_hsm","species","mod")],by="species") |> 
  mutate(hsm=psi-t_hsm*1000-px.mu*1000,
         fsm=frost-t_fsm-lt50.mean,
         margin.sm=case_when(mod%in%c("2sm","hsm")&hsm<0~"drought",
                             mod%in%c("2sm","fsm")&fsm<0~"frost",
                             mod%in%c("2sm")&hsm<0&fsm<0~"both",
                             mod%in%c("2sm")&hsm>0&fsm>0~"core",
                             mod%in%c("hsm")&hsm>0~"core",
                             mod%in%c("fsm")&fsm>0~"core"))

# graph densité fécondité corrigée a chaque marge 
fec.eu_sfm |> 
  filter(!is.na(margin.sm)) |> 
  filter(ISP!=0) |>
  ggplot()+
  # geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
  geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin.sm))+
  facet_wrap(~species,scales="free")+
  scale_x_log10()+
  theme_minimal()

 fec.eu_sfm |> 
  filter(!is.na(margin.sm)) |> 
  filter(ISP!=0) |>
  group_by(species) |> 
  mutate(hsm.core=hsm>quantile(hsm,probs=0.8,na.rm=TRUE),
         fsm.core=fsm>quantile(fsm,probs=0.8,na.rm=TRUE)) |> 
  filter(!(margin.sm=="core"&hsm.core==FALSE&fsm.core==FALSE)) |> 
  ggplot()+
  # geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
  geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin.sm))+
  facet_wrap(~species,scales="free")+
  scale_x_log10()+
  theme_minimal()

# graph densité fécondité non corrigée a chaque marge 
fec.eu_sfm |> 
  filter(!is.na(margin.sm)) |> 
  filter(ISP!=0) |>
  ggplot()+
  geom_boxplot(aes(margin.sm,ISP,color=margin.sm),outlier.colour = NA)+
  # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_y_log10()+
  theme_minimal()


fec.eu_sfm |> 
  filter(!is.na(margin.sm)) |> 
  filter(ISP!=0) |> 
  group_by(species,margin.sm) |> summarise(n=n())
```

# Temporal variability of fecundity

```{r}
files=list.files("data/europe/")[grepl("fecGmMu-",list.files("data/europe/"))]

df.temporal= fecundity.eu_clim |> 
  select(plot,species,BA,fecGmMu,fecGmSd,lon,lat,CV,mat,tmin,map,sgdd,margin)
for (f in files){
  load(paste0("data/europe/",f))
  year=substr(f,9,12)
  mu=summary |> 
    as.data.frame() |> 
    rownames_to_column(var="plot") |> 
    select(plot,matches(unique(fecundity.eu_clim$species))) |> 
    pivot_longer(cols=-plot,
                 names_to = "species",
                 values_to = paste0("fecGmMu_",year))
  rm(summary)
  load(paste0("data/europe/fecGmSd-",year,".rdata"))
  sd=summary |> 
    as.data.frame() |> 
    rownames_to_column(var="plot") |> 
    select(plot,matches(unique(fecundity.eu_clim$species))) |> 
    pivot_longer(cols=-plot,
                 names_to = "species",
                 values_to = paste0("fecGmSd_",year))
  df.temporal <- df.temporal |> 
    left_join(cbind(mu,sd[3]),by=c("plot","species"))
  rm(summary,mu,sd,year)
}

df.temporal2 <- df.temporal|> 
  filter(!is.na(margin)) |> 
  pivot_longer(cols=matches("fecGmMu_"),
               names_to=c("var","year"),
               names_sep = "_",
               values_to="fecYear") |> 
  pivot_longer(cols=matches("fecGmSd_"),
               names_to=c("var2","year2"),
               names_sep = "_",
               values_to="fecSdYear") |> 
  filter(year==year2) |> select(-year2)

df.temporal2 |> 
  mutate(fecYear=na_if(fecYear,0),
         fecSdYear=na_if(fecSdYear,0),
         feccvyear=fecYear/fecSdYear) |> 
  filter(!is.na(fecYear)) |> filter(!is.na(fecSdYear)) |> 
  group_by(margin,plot,species) |> 
  summarise(n_plot=n(),
         fecGmMu_plot=weighted.mean(fecYear,w=1/feccvyear,na.rm=TRUE), # perform weighted mean of the tree on the plot
         se_plot=sqrt(sum(fecSdYear^2)/n_plot),
         cv_plot=se_plot/fecGmMu_plot) |> 
  ungroup() |> 
  ggplot()+
  geom_boxplot(aes(margin,cv_plot,color=margin),outlier.colour = NA)+
  # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_y_log10()+
  theme_minimal()

df.temporal2 |> 
  mutate(fecYear=na_if(fecYear,0),
         fecSdYear=na_if(fecSdYear,0),
         feccvyear=fecYear/fecSdYear) |> 
  filter(!is.na(fecYear)) |> filter(!is.na(fecSdYear)) |> 
  group_by(margin,plot,species) |> 
  summarise(n_plot=n(),
         fecGmMu_plot=weighted.mean(fecYear,w=1/feccvyear,na.rm=TRUE), # perform weighted mean of the tree on the plot
         se_plot=sqrt(sum(fecSdYear^2)/n_plot),
         cv_plot=se_plot/fecGmMu_plot) |> 
  ungroup() |> 
  ggplot()+
  geom_boxplot(aes(margin,fecGmMu_plot,weight=cv_plot,color=margin),outlier.colour = NA)+
  # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_y_log10()+
  theme_minimal()

# both

# df.temporal2 |> 
#   mutate(fecYear=na_if(fecYear,0),
#          fecSdYear=na_if(fecSdYear,0),
#          feccvyear=fecYear/fecSdYear) |> 
#   filter(!is.na(fecYear)) |> filter(!is.na(fecSdYear)) |> 
#   group_by(margin,plot,species) |> 
#   summarise(n_plot=n(),
#          fecGmMu_plot=weighted.mean(fecYear,w=1/feccvyear,na.rm=TRUE), # perform weighted mean of the tree on the plot
#          se_plot=sqrt(sum(fecSdYear^2)/n_plot),
#          cv_plot=se_plot/fecGmMu_plot) |> 
#   ungroup() |> 
#   pivot_longer(cols=c("fecGmMu_plot","cv_plot"),
#              names_to="var",
#              values_to = "val") |> 
#   group_by(species,var) |> mutate(val=scale(val)) |> 
#   ggplot()+
#   geom_boxplot(aes(margin,val,color=var),outlier.colour = NA)+
#   # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
#   facet_wrap(~species,scales="free")+
#   scale_y_log10()+
#   theme_minimal()
 
```

# Improvement :
- weight by uncertainty
- sensitivity to quantiles delineation
- plot margins


## Compute statistics of presence per specie
```{r}
fecundity.eu_clim |> 
  filter(species %in% sp.select.eu) |> 
  filter(BA!=0) |> 
  group_by(species) |> 
  summarise(n=n()) |> arrange(n)
```
## Bootstrap like
Idée : pour chaque espèce, sampling aléatoire d'un nombre n d'individus (ou d'un pourcentage du dataset), petit. Extraction des valeurs de ISP associés au quant975, 025 et 50 de PC1

```{r eval=FALSE, include=FALSE}
df.raw<-data.frame(species=NA,
                   BA=NA,
                   fecGmMu=NA,
                   fecGmSd=NA,
                   name=NA)

for (i in 1:500){
  print(i)
  try<-fecundity |> 
    select(plot,species,BA,fecGmMu,fecGmSd,lon,lat,PC1) |> 
    filter(BA!=0) |> 
    group_by(species) |> 
    sample_frac(0.1) |>
    mutate(quant95=abs(PC1-quantile(PC1,probs=0.95,na.rm=TRUE)[[1]]),
           quant5=abs(PC1-quantile(PC1,probs=0.5,na.rm=TRUE)[[1]]),
           quant05=abs(PC1-quantile(PC1,probs=0.05,na.rm=TRUE)[[1]])) |>
    pivot_longer(cols=c("quant95","quant5","quant05")) |> 
    ungroup() |> group_by(species,name) |> slice(which.min(value)) |> 
    select(species,BA,fecGmMu,fecGmSd,name)
  df.raw=rbind(df.raw,try)
}

df.raw |> 
  filter(!is.na(species)) |> 
  mutate(fecGmCv=100*fecGmSd/fecGmMu,
         ISP=fecGmMu/BA) |> 
  ggplot()+
  geom_density(aes(x=ISP,y=..density..,weight=fecGmCv,color=name))+
  facet_wrap(~species,scales="free")+
  scale_x_log10()+
  theme_minimal()

  
```

## sensitivity to quantiles
```{r eval=FALSE, include=FALSE}
prob.low=seq(0.01,0.47,0.01)
prob.high=seq(0.99,0.53,-0.01)
sensibility<-data.frame(species=unique(fecundity$species),
                        low.mid=NA,
                        low.high=NA,
                        high.mid=NA,
                        low.mid.sd=NA,
                        low.high.sd=NA,
                        high.mid.sd=NA) |> crossing(i=1:47)
for (i in 1:47){
  print(i)
  clim_i=BA_select |> 
    filter(!species%in%c("sorbusAucupari","sorbusAucuparia","cedrusAtlantic")) |> 
    filter(BA!=0) |> 
    group_by(species) |> 
    summarise(quant.high=quantile(PC1,probs=prob.high[i],na.rm=TRUE)[[1]],
              quant525=quantile(PC1,probs=0.525,na.rm=TRUE)[[1]],
              quant475=quantile(PC1,probs=0.475,na.rm=TRUE)[[1]],
              quant.low=quantile(PC1,probs=prob.low[i],na.rm=TRUE)[[1]])
  fecundity_i<-fecundity |> 
    select(plot,species,fecGmMu,fecGmSd,PC1,BA) |> 
    left_join(clim_i)|> 
    mutate(fecGmCv=100*fecGmSd/fecGmMu,
           ISP=fecGmMu/BA,
           margin=case_when(PC1<quant.low~"low",
                          PC1>quant475&PC1<quant525~"opt",
                          PC1>quant.high~"up",
                          TRUE~NA)) |> 
    filter(!is.na(margin)) |>
    filter(ISP!=0 | BA !=0) |>
    filter(!is.nan(fecGmCv))
  for (sp in unique(clim_i$species)){
    fecundity_i |> filter(species==sp)->test
    sensibility[sensibility$species==sp&sensibility$i==i,c("low.mid","low.high","high.mid")]=
      list(
        t.test(test[test$margin=="low","ISP"],test[test$margin=="opt","ISP"])$p.value,
        t.test(test[test$margin=="low","ISP"],test[test$margin=="up","ISP"])$p.value,
        t.test(test[test$margin=="up","ISP"],test[test$margin=="opt","ISP"])$p.value
      )
  }
  }
sensibility |> 
  pivot_longer(cols=c("low.mid","low.high","high.mid")) |> 
  ggplot(aes(i,value,color=name))+
    geom_line()+
    facet_wrap(~species)
```

# Annexe 


Check data for steph
```{r eval=FALSE, include=FALSE}
plotData |> 
  filter(lat>48&lat<50) |> 
  filter(lon<18&lon>12) -> czplot

cztree<- as.data.frame(BA)|> 
  rownames_to_column(var="plot") |> 
  filter(plot %in% czplot$plot) |> 
  pivot_longer(cols=-plot,
               names_to = "species",
               values_to = "BA") |> 
  filter(BA>0)

cztree |> 
  group_by(plot,species) |> summarise(n=n())
cztree |> 
  # filter(plot%in% mono$plot) |>  
  group_by(species) |> 
  summarise(n=n()) |> arrange(n)
cztree |> select(plot,species) |> distinct() |> group_by(plot) |> summarise(n=n()) |> filter(n<2)->mono
```