---
title: "Bristish Ecological Society - presentation"
author: "Anne Baranger"
date: "2023-12-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE,message=FALSE)
library(targets)
lapply(c("stringr","tidyr","dplyr","tibble", # data analysis
         "ggplot2","viridis", # plot
         "factoextra", "nlme","modi","rstan","shinystan",# stat & models
         "rworldmap" # geographic data #terre, sf
         ),require,character.only=TRUE)
theme_set(theme_minimal()) # ggplot theme


# load necessary data #

## mastif data ##
# tar_load(mastif.eu,store="target_data")
# tar_load(mastif.am,store="target_data")

## gbif data ##
#tar_load(meanClimate_species,store="target_gbif")


## results from selection ##
# to be nuanced 
tar_load(species_selection,store="target_nfi")

## species selection ##
tar_load(phylo.select,store="target_nfi")

tar_load(sp.select.eu,store="target_nfi")
tar_load(sp.select.am,store="target_nfi")
tar_load(species_class,store="target_nfi")
phylo.select<-phylo.select |> left_join(species_class) |> 
  mutate(zone=case_when(species=="abiesCephalon"~"europe",
                        species=="acerCircinat"~"west",
                        species=="acerSpicatum"~"east",
                        species=="amelanchArborea"~"east",
                        species=="asiminaTriloba"~"east",
                        species=="cedrusAtlantic"~"europe",
                        species=="frangulaAlnus"~"europe",
                        species=="linderaBenzoin"~"east",
                        TRUE~zone)) 
  


# small format of gbif
df.gbif.short <- species_selection$df.gbif |>
  separate(clim,into = c("clim","range")) |> 
  pivot_wider(names_from = range,
                values_from = clim_niche)

#compute climate range
# df.range<-species_selection$df.gbif |> 
#     dplyr::select(species,clim,clim_niche) |> 
#     filter(grepl("range",clim)) |> 
#     separate(clim,into=c("var","drop")) |> 
#     dplyr::select(-drop) |> 
#     rename(range="clim_niche")

## fecundity data ##
tar_load(fecundity.eu_clim,store="target_nfi")
tar_load(fecundity.am_clim,store="target_nfi")

worldmap <- sf::st_as_sf(getMap(resolution = "high"))
```

# Associate species to biomes
```{r}
fec_tot=rbind(fecundity.eu_clim |> mutate(block="europe"),
                fecundity.am_clim|> mutate(block="america")) |>
  left_join(phylo.select) |> 
  dplyr::select(plot,lon,lat,species,BA,fecGmMu,fecGmSd,sgdd,wai,pet,map,mat,block,zone) |> 
  filter(BA!=0) |> #rm absences
  mutate(dh=12*pet-map)
ecoregions=sf::read_sf(dsn="data/WWF/official", #read ecoregions
                   layer="wwf_terr_ecos") |>
  select(BIOME,ECO_NAME,geometry) |> 
  filter(!BIOME%in%c(98,99))
sf::sf_use_s2(FALSE)
points=sf::st_as_sf(fec_tot[,c("lon","lat")],coords=c("lon","lat"),crs=sf::st_crs(ecoregions))
points_biome=sf::st_join(points,ecoregions) 

cbind(fec_tot,
      biome=points_biome$BIOME) |> 
  group_by(species) |> mutate(n_sp=n()) |> 
  group_by(n_sp,species,biome) |>
  summarise(n=n()/n_sp) |> unique() |> ungroup() |> 
  group_by(species) |> 
  arrange(desc(n), .by_group = TRUE) %>%
  slice_head(n = 2) |> 
  filter(n>0.1) |> 
  ungroup() |> 
  select(species,biome,n) |> 
  mutate(biome=case_when(biome==13 ~ 12,
                         TRUE ~ biome))->sp_biome
```

# Introduction 

## Map of MASTIF data
```{r}
p<-rbind(readRDS("target_data/objects/mastif.am")$df.tree,
      readRDS("target_data/objects/mastif.eu")$df.tree) |> 
  filter(species %in% c(sp.select.am,sp.select.eu)) |> 
  select(species,lon,lat) |> 
  distinct() |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_jitter(aes(x=lon,y=lat,color=species),size=1.3,alpha=0.5)+
  theme_minimal()+
  xlim(-135,24) + 
  ylim(27,73)+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "none")+
  labs(color="Species")+
  ggtitle(label="Maps of MASTIF observations network for 85 species")
p
# ggsave( p,
#         filename = "bes_fig/MastifPlot.png",
#         dpi = 600,
#         width = 15,
#         height=9,
#         units = "cm")
```

## Species selection
```{r}
species_selection$df.gbif |> 
  # filter(species %in% sp.select.am) |> 
  filter(grepl("mat",clim)) |> 
  pivot_wider(names_from = clim,
              values_from = clim_niche) |> 
  rename(quant=mat.low,rank=mat.rlow) |> 
  pivot_longer(cols=c("quant","rank"),
               names_to = "low_type",
               values_to = "low") |> 
  rename(quant=mat.high,rank=mat.rhigh) |> 
  pivot_longer(cols=c("quant","rank"),
               names_to = "high_type",
               values_to = "high") |> 
  filter(low_type==high_type) |> 
  mutate(source="gbif") |> 
  select(-species_l)->gbif.sp
gbif.sp |> select(species,mat.opt,mat.range)

species_selection$df.select |> 
  # filter(species %in% sp.select.am) |> 
  select(species,matches("mat")) |> 
  rename(quant=mat.low,rank=mat.rlow) |> 
  pivot_longer(cols=c("quant","rank"),
               names_to = "low_type",
               values_to = "low") |> 
  rename(quant=mat.high,rank=mat.rhigh) |> 
  pivot_longer(cols=c("quant","rank"),
               names_to = "high_type",
               values_to = "high") |> 
  filter(low_type==high_type) |> 
  mutate(source="mastif")->mastif.sp

df.plot=rbind(mastif.sp,
      gbif.sp |> select(-mat.opt,-mat.range)) |> 
  # filter(species %in% sample(unique(species_selection$df.gbif$species),15)) |> 
  left_join(gbif.sp |> select(species,mat.opt) |> unique()) |> arrange(species) |> 
  left_join(species_selection$df.select |> select(species,select.quant,select.rank)) |> 
  filter(!is.na(mat.opt)) |> 
  # rename(quant=select.quant,rank=select.rank) |>
  # pivot_longer(cols=c("quant","rank"),
  #              names_to = "select_type",
  #              values_to = "select") |>
  # filter(select_type==high_type) |>
  mutate(species=forcats::fct_reorder(species,mat.opt)) 
p<-df.plot |> 
  filter(species %in% sample(unique(df.plot$species),25)) |> 
  filter(!species %in% c("sorbusAmerican","corylusCornuta","acerSpicatum")) |> 
  filter(high_type=="quant") |> 
  mutate(select.quant=if_else(select.quant==TRUE,"Selected","Not selected")) |> 
  ggplot(aes(x=(low+high)/2,y=species))+
  geom_pointrange(aes(xmin=low,xmax=high,color=source),position=position_dodge(width = 0.5))+
  facet_wrap(~select.quant,scales = "free_y")+
  scale_color_manual(values=c("grey","black"))+
  labs(color="Source",
       x="Temperature range")
p
# ggsave( p,
#         filename = "bes_fig/SpeciesSelect.png",
#         dpi = 600,
#         width = 15,
#         height=9,
#         units = "cm")
```

## Map of NFI plots


### Color by climate
```{r plots10}
# plots of selected species
## mat 
fecundity.fit.1 |> 
  dplyr::select(plot,lat,lon,dh,mat) |> 
  unique() |> 
  sample_frac(0.1) |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=mat))+
  scale_color_gradient2( low = "steelblue1",mid="bisque", high = "firebrick1")+
  xlim(-135,24) + 
  ylim(27,73)+
  theme_minimal()+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom")+
  ggtitle("10 % of NFI plots")

## dh
fecundity.fit.1|> 
  filter(dh>=(-2000)) |> 
  dplyr::select(plot,lat,lon,dh,mat) |> 
  unique() |> 
  sample_frac(0.1) |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=dh),size=0.4)+
  scale_color_gradient2( low = "steelblue1",mid="bisque", high = "firebrick1",breaks=c(-1000,0,1000))+
  xlim(-135,24) + 
  ylim(27,73)+
  theme_minimal()+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom")+
  ggtitle("10 % of NFI plots")
```

### Color by abundance and biogeography 

```{r plots10}
# abundance of nfi plots
p<-fecundity.fit.1 |> 
  filter(dh>=(-2000)) |> 
  dplyr::select(plot,lat,lon,dh,mat) |> 
  unique() |> 
  # sample_frac(0.1) |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_hex(aes(x=lon,y=lat),bins=80)+
  xlim(-135,24) + 
  ylim(27,73)+
  theme_minimal()+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom")+
  ggtitle("10 % of NFI plots")+
  scale_fill_continuous(type="viridis",trans="sqrt",breaks=c(1,100,1000,5000))
# ggsave( p,
#         filename = "bes_fig/NFIPlot.png",
#         dpi = 600,
#         width = 15,
#         height=9,
#         units = "cm")


# plots by biogeographical zone
p<-fecundity.fit.1 |> 
  filter(dh>=(-2000)) |> 
  dplyr::select(species,lat,lon,dh,mat) |> 
  unique() |>
  mutate(zone=factor(zone,levels=c("west","east","europe"))) |> 
  sample_frac(0.1) |>
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=zone))+#,bins=80
  xlim(-135,24) + 
  ylim(27,73)+
  theme_minimal()+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom")+
  ggtitle("10 % of NFI plots")+
  scale_fill_continuous(type="viridis",trans="sqrt",breaks=c(1,100,1000,5000))+
  scale_color_manual(values=c("firebrick4","firebrick1","slateblue"))+
  labs(color="Biogeographical zone")
# ggsave( p,
#         filename = "bes_fig/BiogeoPlot.png",
#         dpi = 600,
#         width = 15,
#         height=9,
#         units = "cm")
```


## Geographical zone

### West species

```{r mapspw}
fecundity.fit.1 |> 
  filter(dh>=(-2000)) |> 
  filter(zone=="west") |> 
  sample_frac(0.1) |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=species),size=0.4)+
  xlim(-135,-90) + 
  ylim(27,73)+
  theme_minimal()+
  stat_ellipse(aes(x=lon,y=lat,color=species))
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom")+
  ggtitle("10 % of NFI plots")
  
 
```


# Does temperature limit the fecundity in species' margins?

## Map of margins and fecundity boxplot for 3 european species
```{r}
species.sub<- fecundity.fit.1 |> 
  filter(species %in% c("piceaAbies","fagusSylvatic","quercusIlex")) |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot")),
         species=factor(species,levels=c("piceaAbies","fagusSylvatic","quercusIlex"))
         ) |> 
  filter(!is.na(margin.temp)) |> #&!is.na(margin.deficit)
  ungroup() 

map<-species.sub |> 
  ggplot()+
  geom_point(aes(x=lon,y=lat,color=margin.temp))+
  geom_sf(data=worldmap,fill=NA)+
  xlim(c(-10,35))+
  ylim(c(35,70))+
  facet_wrap(~species)+
  labs(color="Temperature margin:")+
  scale_color_manual(values=c("steelblue1","bisque2","firebrick1"))+
  theme(axis.title = element_blank(),
        legend.position = "top")

box<-species.sub |> 
  ggplot(aes(margin.temp,log(ISP),fill=margin.temp))+
  geom_boxplot(outlier.colour = NA)+
  scale_fill_manual(values=c("steelblue1","bisque2","firebrick1"))+
  facet_wrap(~species)+
  theme(legend.position = "none")+
  labs(y="Fecundity (log)",
       x="")
cowplot::plot_grid(map,box,nrow=2 )
# ggsave( map,
#         filename = "bes_fig/MapMargins.png",
#         dpi = 600,
#         width = 15.5,
#         height=10,
#         units = "cm")
```
- mettre une carte de l'am du n pour faire plaisir a g

## Analysis per continent

```{r}
rm(map,box,species.sub)

tar_load(fitcontinentdiscrete,store="target_fit")
fitcontinentdiscrete |>
  ggplot(aes(margin.temp,posterior,color=zone))+
  geom_boxplot(outlier.colour = NA)+
  scale_color_manual(values=c("firebrick4","firebrick1","slateblue"))+
  theme(axis.title=element_blank(),
        legend.position = "bottom")+
  geom_hline(yintercept = 0,color="gray12",type="dashed")+
  labs(title="Log ratio of fecundity relative to \nthe mean fecundity of the niche",
       color="Geographical \nzone:")-> p

p
# ggsave( p,
#        filename = "bes_fig/MarginTempCont.png",
#        dpi = 600,
#        width = 18,
#        height=14,
#        units = "cm")
```

## Analysis per continent & taxa

```{r conttaxa}
sp.bayes=fecundity.fit.1 |> 
  filter(!is.na(margin.temp)) 
  # filter(species %in% select.species[3])

# mid_isp<-sp.bayes |> 
#     group_by(species,margin.temp) |> 
#     summarise(mid_isp=median(ISP)[[1]]) |> 
#     filter(margin.temp=="midtemp") |>
#     ungroup() |> 
#     dplyr::select(species,mid_isp)
mid_isp<-sp.bayes |> 
    group_by(species) |> 
    summarise(mid_isp=median(ISP)[[1]]) |> 
    ungroup() |> 
    dplyr::select(species,mid_isp)
sp.bayes<-sp.bayes |>
  left_join(mid_isp) |>
  mutate(dISP=ISP/mid_isp,
         s_ISP=s_ISP/(mid_isp^2)) |> 
  filter(!is.na(dISP))

## angios
sp.taxa=sp.bayes |> 
  filter(taxa=="angiosperm")
X=model.matrix(~margin.temp*block, sp.taxa)

if(file.exists("fit.allsp.margintemp.angio.RData")){
  load("fit.allsp.margintemp.angio.RData")
}else{
  data_list=list(N=dim(sp.taxa)[1],
               NX=ncol(X),
               S=nlevels(as.factor(sp.taxa$species)),
               species=as.numeric(as.factor(sp.taxa$species)),
               ISP=sp.taxa$dISP,
               X=X)
               
  fit.lm <- stan(file = "lmm_dif.stan",
                    data=data_list,
                    iter=1000,
                    chains=3,
                    core=3)
  save(fit.lm,file="fit.allsp.margintemp.angio.RData")

}

# launch_shinystan(fit.lm)
posteriors_fec<-as.data.frame(fit.lm) |> 
  dplyr::select(!matches("beta_")) |> 
  dplyr::select(matches("beta"))
colnames(posteriors_fec)[1:ncol(X)]=colnames(X)


# contrast.max=
contrast_cor=as.data.frame(X) |> 
  unique() |>
  mutate(margin.temp=case_when(margin.tempcold==1~"cold",
                               margin.temphot==1~"hot",
                               TRUE~"mid"),
         zone=case_when(blockeurope==1~"europe",
                        TRUE~"america")) |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","mid","hot"))) |> 
  tibble::rownames_to_column(var = "code") |> 
  dplyr::select(code,margin.temp,zone) |> 
  mutate(contrast=paste0(margin.temp,"_",zone))

contrast_mat= X |> unique() |>  t()
contrast_post=as.matrix(posteriors_fec)[,1:ncol(X)] %*% contrast_mat

as.data.frame(contrast_post) |> 
  pivot_longer(cols=everything(),
               names_to = "code") |>
  left_join(contrast_cor,by="code") |> 
  mutate(zone=factor(zone,levels=c("america","europe"))) |> 
  ggplot(aes(margin.temp,value,color=zone))+
  geom_boxplot(outlier.colour = NA)+
  scale_color_manual(values=c("firebrick","slateblue"))+
  theme(axis.title=element_blank(),
        legend.position = "bottom")+
  geom_hline(yintercept = 0,color="gray12",type="dashed")+
  labs(title="Log ratio of fecundity relative to \nthe mean fecundity of the niche, \n Angiosperms",
       color="Geographical zone:")-> p

p
ggsave( p,
        filename = "bes_fig/MarginTempCont_angios.png",
        dpi = 600,
        width = 18,
        height=14,
        units = "cm")

## gymno
sp.taxa=sp.bayes |> 
  filter(taxa=="gymnosperm")
X=model.matrix(~margin.temp*zone, sp.taxa)

if(file.exists("fit.allsp.margintemp.gymno.RData")){
  load("fit.allsp.margintemp.gymno.RData")
}else{
  data_list=list(N=dim(sp.taxa)[1],
               NX=ncol(X),
               S=nlevels(as.factor(sp.taxa$species)),
               species=as.numeric(as.factor(sp.taxa$species)),
               ISP=sp.taxa$dISP,
               X=X)
               
  fit.lm <- stan(file = "lmm_dif.stan",
                    data=data_list,
                    iter=1000,
                    chains=3,
                    core=3)
  save(fit.lm,file="fit.allsp.margintemp.gymno.RData")

}

# launch_shinystan(fit.lm)
posteriors_fec<-as.data.frame(fit.lm) |> 
  dplyr::select(!matches("beta_")) |> 
  dplyr::select(matches("beta"))
colnames(posteriors_fec)[1:ncol(X)]=colnames(X)


contrast_cor=as.data.frame(X) |> 
  unique() |>
  mutate(margin.temp=case_when(margin.tempcold==1~"cold",
                               margin.temphot==1~"hot",
                               TRUE~"mid"),
         zone=case_when(zoneeurope==1~"europe",
                        zonewest==1~"west",
                        TRUE~"east")) |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","mid","hot"))) |> 
  tibble::rownames_to_column(var = "code") |> 
  dplyr::select(code,margin.temp,zone) |> 
  mutate(contrast=paste0(margin.temp,"_",zone))

contrast_mat= X |> unique() |>  t()
contrast_post=as.matrix(posteriors_fec)[,1:ncol(X)] %*% contrast_mat

as.data.frame(contrast_post) |> 
  pivot_longer(cols=everything(),
               names_to = "code") |>
  left_join(contrast_cor,by="code") |> 
  mutate(zone=factor(zone,levels=c("west","east","europe"))) |> 
  ggplot(aes(margin.temp,value,color=zone))+
  geom_boxplot(outlier.colour = NA)+
  scale_color_manual(values=c("firebrick4","firebrick1","slateblue"))+
  theme(axis.title=element_blank(),
        legend.position = "bottom")+
  geom_hline(yintercept = 0,color="gray12",type="dashed")+
  labs(title="Log ratio of fecundity relative to \nthe mean fecundity of the niche, \nGymnosperms",
       color="Geographical \nzone:")-> p

p
ggsave( p,
        filename = "bes_fig/MarginTempCont_gymno.png",
        dpi = 600,
        width = 18,
        height=14,
        units = "cm")
```

# Does the structure of the climatic space impact the effect of temperature on fecundity?

## Climatic space and correlation of DH/mat inside species distribution

### ACP of climatic space
```{r climacp, eval=FALSE, fig.height=12, fig.width=19, include=FALSE}
z="europe"
for (z in c("west","east","europe")){
  clim.acp=fecundity.fit.1|> 
    filter(zone==z) |> 
    # sample_n(size = 10000) |>
    # filter(ISP<quantile(ISP,probs=0.97)) |>  # remove extreme values
    dplyr::select(plot,zone,taxa,lon,lat,mat,tmin,map,pmax,pet) |> #,cmi_min,sgdd,cmi_max,wai,cmi_mean
    filter(dh>=(-2000)) |> 
    unique() |> 
    drop_na()
  acp=prcomp(clim.acp[,c("mat","dh")],scale=TRUE)
# fviz_pca_ind(acp)
# fviz_pca_var(acp,axes=c(1,3))
  species=fecundity.fit.1 |>  
    filter(zone==z) |> 
    # sample_n(size = 10000) |>
    # filter(ISP<quantile(ISP,probs=0.97)) |>  # remove extreme values
    dplyr::select(species,mat,map,pet) |> #,cmi_min,sgdd,cmi_max,wai,cmi_mean
    filter(dh>=(-2000)) |> 
    group_by(species) |> 
    summarise(mat=median(mat),
              dh=median(dh))
  
  pred=cbind(species,
             predict(acp,species))

  pca.plot=fviz_pca(acp,
           axes=c(1,2),
           # geom.ind="point",
           geom.var=c("arrow","text"),
           # palette=c("firebrick1","slateblue","firebrick4"),
           # habillage=clim.acp$species,
           col.var="black",
           col.ind=NA,
           alpha.ind=0.05,
           # addEllipses=TRUE,
           )+
    theme(legend.position = "none")+
    geom_point(data=as.data.frame(pred),aes(x=PC1,y=PC2,color=species))+
    ggtitle( label = str_to_title(z))+
    theme(plot.title = element_text(hjust=0.48))
    # labs(color="Geographical \nzone",
         # fill="Geographical \nzone",
         # shape="Geographical \nzone")
 print(pca.plot)
 
 cor<-fecundity.fit.1|> 
    filter(zone==z) |> 
    # sample_n(size = 10000) |>
    # filter(ISP<quantile(ISP,probs=0.97)) |>  # remove extreme values
    dplyr::select(plot,zone,species,taxa,lon,lat,mat,tmin,map,pmax,pet) |> #,cmi_min,sgdd,cmi_max,wai,cmi_mean
    filter(dh>=(-2000)) |> 
    group_by(species) |> 
    summarise(cormatdh=cor(mat,dh,use="complete")) |> 
   ggplot(aes(cormatdh))+
   geom_histogram()+
   xlim(c(-1,1))+
   ylim(c(0,9))+
   geom_vline(xintercept = 0,color="darkred",linetype="dashed")+
   theme(axis.title = element_blank())
 
 
 assign(paste0(z,"_plot"),
        cowplot::plot_grid(pca.plot,cor,nrow=2,rel_widths = c(1,1),rel_heights = c(2,1)),
        env=.GlobalEnv)
}

cowplot::plot_grid(west_plot,east_plot,europe_plot,ncol = 3)->p
# ggsave( p,
#         filename = "bes_fig/AcpClim.png",
#         dpi = 600,
#         width = 19,
#         height=12,
#         units = "cm")
z
pca.plot=fviz_pca(acp,
           axes=c(1,2),
           geom.ind="point",
           geom.var=c("arrow","text"),
           # palette=c("firebrick1","slateblue","firebrick4"),
           # habillage=clim.acp$species,
           col.var="black",
           # col.ind=NA,
           alpha.ind=0.05,
           # addEllipses=TRUE,
           )
```

### Correlation of DH/MAT

```{r cordhmat}
fecundity.fit.1 |> 
  group_by(species) |> 
  mutate(cor.mat.dh=cor(dh,mat,use="complete.obs")) |> 
  dplyr::select(species,cor.mat.dh,zone) |> 
  unique()|> 
  ggplot(aes(cor.mat.dh,color=zone))+
  geom_density()+
  scale_color_manual(values=c("firebrick1","slateblue","firebrick4"))+
  labs(color="Geographical \nzone",
       y="",
       x="Intraspecific correlation between MAT and DH")
```

### Exemple of highly correlated species

```{r corex,fig.height=6,fig.width=12}
fecundity.fit.1 |> 
  filter(species %in% c("caryaOvata","pinusEdulis")) |> 
  filter(dh>quantile(dh,probs = 0.01,na.rm=TRUE)) |> filter(dh<quantile(dh,probs=0.99,na.rm=TRUE)) |> 
  sample_n(max(n())) |> 
  ggplot(aes(mat,dh,color=margin.temp))+
  geom_point()+
  geom_smooth(method="lm",aes(color=NULL))+
  facet_wrap(~species)+
  scale_color_manual(values=c("bisque2","steelblue1","firebrick1"))+
  labs(color="Temperature margins :",
       x="Mean annual temperature (°C)",
       y="Water deficit (mm)")->p
p
# ggsave( p,
#         filename = "bes_fig/SpCorEx.png",
#         dpi = 600,
#         width = 13,
#         height=7,
#         units = "cm")
rm(p)
```

## Effect of aridity gradient in species margins on fecundity, for species with high correlation between mat & dh

```{r ariditycor}
tar_load(fitcontinentdiscreteexcl,store="target_fit")
fitcontinentdiscreteexcl |> 
  ggplot(aes(margin.temp,posterior,color=zone))+
  geom_boxplot(outlier.colour = NA)+
  scale_color_manual(values=c("firebrick4","firebrick1","slateblue"))+
  theme(axis.title=element_blank(),
        legend.position = "bottom")+
  geom_hline(yintercept = 0,color="gray12")+
  labs(title="Log ratio of fecundity relative to \nthe mean fecundity of the niche",
       color="Geographical \nzone:")-> p
p
ggsave( p,
        filename = "bes_fig/MarginTempCont_spcor70.png",
        dpi = 600,
        width = 18,
        height=14,
        units = "cm")
```


## Effect of aridity gradient in species margins on fecundity, for species with low correlation between mat & dh

```{r ariditygrad}
tar_load(fitcontinentcontinous)
out_ancova_cont=fitcontinentcontinous$posterior
out_ancova_cont |> 
  mutate(zone=factor(str_to_title(zone),levels=c("West","East","Europe")),
         margin.temp=case_when(margin.temp=="midtemp"~"center",
                               TRUE~margin.temp)) |> 
  ggplot()+
  geom_point(aes(dh,log(dISP)-mean,color=margin.temp),size=0.6,alpha=0.05)+
  geom_ribbon(aes(x=dh,ymin=quant05,ymax=quant95,fill=margin.temp),alpha=0.8)+
  geom_line(aes(dh,fit,group=margin.temp),size=0.2,color="black")+
  geom_boxplot(aes(dh,y=-12,color=margin.temp),position=position_dodge(width=2),outlier.colour = NA)+
  scale_fill_manual(values=c("bisque2","steelblue1","firebrick1"))+
  scale_color_manual(values=c("bisque2","steelblue1","firebrick1"))+
  facet_wrap(~zone)+
  theme(text = element_text(size=13),
        legend.position = "bottom")+
  labs(y="Log ratio of fecundity relative to the mean",
       x="Water deficit (mm, centered by species)",
       color="Temperature margin",
       fill="Temperature margin")->p
p


out_ancova_cont |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot"))) |> 
  ggplot()+
  geom_point(aes(dh,log(dISP)-mean,color=zone),size=0.6,alpha=0.05)+
  geom_ribbon(aes(x=dh,ymin=quant05,ymax=quant95,fill=zone),alpha=0.8)+
  geom_line(aes(dh,fit,group=zone),size=0.2,color="black")+
  geom_boxplot(aes(dh,y=-10,color=zone),position=position_dodge(width=1),outlier.colour = NA)+
  geom_boxplot(aes(x=4,y=log(dISP)-mean,color=zone),position=position_dodge(width=1),outlier.colour = NA)+
  scale_fill_manual(values=c("firebrick1","slateblue","firebrick4"))+
  scale_color_manual(values=c("firebrick1","slateblue","firebrick4"))+
  theme(axis.text.x = element_text(colour = inferno(6)))+
  facet_wrap(~margin.temp)


out_ancova_cont |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot"))) |> 
  filter(margin.temp=="hot") |> 
  ggplot()+
  geom_point(aes(dh,log(dISP)-mean,color=zone),size=0.6,alpha=0.01)+
  geom_ribbon(aes(x=dh,ymin=quant05,ymax=quant95,fill=zone),alpha=0.8)+
  geom_line(aes(dh,fit,color=zone),size=0.2)+
  geom_boxplot(aes(dh,y=-10,color=zone),position=position_dodge(width=1),outlier.colour = NA)+
  # geom_boxplot(aes(x=4,y=log(dISP)-mean,color=zone),position=position_dodge(width=1),outlier.colour = NA)+
  scale_fill_manual(values=c("firebrick1","slateblue","firebrick4"))+
  scale_color_manual(values=c("firebrick1","slateblue","firebrick4"))+
  theme(axis.text.x = element_text(colour = inferno(6)),
        text = element_text(size=13))+
  ggtitle("Fecundity variation with aridity in species hot margin")+
  labs(y="Log ratio of fecundity relative to the mean",
       x="Water deficit (mm, centered by species)",
       color="Biogeographical zone",
       fill="Biogeographical zone")->p

out_ancova_cont |> 
  mutate(zone=factor(zone,levels=c("west","east","europe")),
         margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot"))) |>
  ggplot(aes(margin.temp,fit,color=zone))+
  geom_boxplot(outlier.color = NA)+
  scale_color_manual(values=c("firebrick4","firebrick1","slateblue"))

# ggsave( p,
#         filename = "bes_fig/Int_3.png",
#         dpi = 600,
#         width = 17,
#         height=13,
#         units = "cm")
```


## Effect of aridity gradient in species margins on fecundity, difference between taxa, for species with low correlation between mat & dh

```{r ariditygrad}
sp.ancova=rbind(fecundity.eu_clim |> mutate(block="europe"),
                fecundity.am_clim|> mutate(block="america")) |>
  left_join(phylo.select) |> 
  dplyr::select(plot,lon,lat,species,taxa,BA,fecGmMu,fecGmSd,sgdd,wai,pet,map,mat,block,zone) |> 
  filter(BA!=0) |> #rm absences
  mutate(dh=12*pet-map) |> 
  filter(dh>(-2000)) |>
  group_by(species) |> 
  # compute weighted quantiles of sgdd and wai, and correlation between wai and sgdd
  mutate(margin.temp=case_when(mat<=weighted.quantile(mat,w=BA, prob=0.1)[[1]]~"cold",
                               mat<weighted.quantile(mat,w=BA, prob=0.55)[[1]]&
                                 mat>weighted.quantile(mat,w=BA, prob=0.45)[[1]]~"midtemp",
                               mat>=weighted.quantile(mat,w=BA, prob=0.9)[[1]]~"hot"),
         margin.temp=factor(margin.temp,levels=c("midtemp","cold","hot"))
         ) |>
  # filter(dh>quantile(dh,probs = 0.01,na.rm=TRUE)) |> filter(dh<quantile(dh,probs=0.99,na.rm=TRUE)) |> 
  filter(!is.na(margin.temp)) |> 
  # filter(!is.na(margin.temp)&!is.na(margin.deficit)) |>
  # ungroup() |> group_by(species,margin.temp) |> 
  mutate(dh_old=dh,
         dh=scale(dh,center=TRUE,scale=FALSE),
         dh2=dh^2) |> 
  ungroup() |> 
  mutate(ISP=fecGmMu/BA,
         s_ISP=fecGmSd/(BA^2))|> 
  filter(!is.na(ISP)) |> 
  filter(!is.na(dh))
mid_isp<-sp.ancova |> 
    group_by(species) |> 
    summarise(mid_isp=median(ISP)[[1]]) |> 
    ungroup() |> 
    dplyr::select(species,mid_isp)
sp.ancova<-sp.ancova |>
  left_join(mid_isp) |>
  mutate(dISP=ISP/mid_isp,
         s_ISP=s_ISP/(mid_isp^2)) |> 
  filter(!is.na(dISP))

folder="mod_ancova_taxa/"
out_ancova_taxa=sp.ancova |> 
  mutate(taxa=NA,
         fit=NA,
         quant05=NA,
         quant95=NA,
         mean=NA) |> 
  slice(0)
g="angiosperm"
for (z in c("europe","america")){
  print(z)
  # filter data
  sp.zone=sp.ancova |> 
    filter(block==z) |> 
    filter(taxa==g)

  X=model.matrix(~margin.temp*dh, sp.zone)
   
  if(file.exists(paste0(folder,"ancova_",z,"_angiosperm_lin.RData"))){
    load(paste0(folder,"ancova_",z,"_angiosperm_lin.RData"))
  }else{
    data_zone=list(N=dim(sp.zone)[1],
                   NX=ncol(X),
                   S=nlevels(as.factor(sp.zone$species)),
                   species=as.numeric(as.factor(sp.zone$species)),
                   ISP=sp.zone$dISP,
                   X=X,
                   NULL)
    fit.ancova_lin <- stan(file = "lmm_ancova.stan",
                       data=data_zone,
                       iter=1000,
                       chains=3,
                       core=3,
                       include=FALSE,
                       pars=c("beta_raw"))
    save(fit.ancova_lin,file=paste0(folder,"ancova_",z,"_angiosperm_lin.RData"))    
  }
  
  fit=fit.ancova_lin
  
  posteriors_fec<-as.data.frame(fit) |>
    dplyr::select(!matches("beta_")) |>
    dplyr::select(matches("beta"))
  colnames(posteriors_fec)[1:ncol(X)]=colnames(X)
  speciescode=sp.zone |> ungroup()|> dplyr::select(species) |> mutate(code=as.numeric(as.factor(species))) |> unique()
  speciesranef=as.data.frame(summary(fit)$summary) |> 
    rownames_to_column(var="parameter") |> 
    filter(grepl("alpha",parameter)) |> 
    dplyr::select(parameter,mean) |> 
    mutate(code=as.numeric(gsub(".*?\\[([^]]*)\\].*", "\\1", parameter))) |> 
    left_join(speciescode) |> 
    dplyr::select(species,mean)
  
  contrast_mat= X |>  t()
  contrast_post=as.matrix(posteriors_fec)[,1:ncol(X)] %*% contrast_mat
  fit=apply(contrast_post,2,median)
  quant05=apply(contrast_post,2,quantile,probs=0.05)
  quant95=apply(contrast_post,2,quantile,probs=0.95)
  
  output=cbind(sp.zone,
               fit=fit,
               quant05=quant05,
               quant95=quant95) |>
    left_join(speciesranef,by="species") 

  out_ancova_taxa=rbind(out_ancova_taxa,
                   output)
  rm(output,quant05,quant95,fit,contrast_post,contrast_mat,speciesranef,
     speciescode,posteriors_fec,data_zone,sp.mod,sp.zone)
}

g="gymnosperm"
for (z in c("europe","east","west")){
  print(z)
  # filter data
  sp.zone=sp.ancova |> 
    filter(zone==z) |> 
    filter(taxa==g)

  X=model.matrix(~margin.temp*dh, sp.zone)
   
  if(file.exists(paste0(folder,"ancova_",z,"_gymnosperm_lin.RData"))){
    load(paste0(folder,"ancova_",z,"_gymnosperm_lin.RData"))
  }else{
    data_zone=list(N=dim(sp.zone)[1],
                   NX=ncol(X),
                   S=nlevels(as.factor(sp.zone$species)),
                   species=as.numeric(as.factor(sp.zone$species)),
                   ISP=sp.zone$dISP,
                   X=X,
                   NULL)
    fit.ancova_lin <- stan(file = "lmm_ancova.stan",
                       data=data_zone,
                       iter=1000,
                       chains=3,
                       core=3,
                       include=FALSE,
                       pars=c("beta_raw"))
    save(fit.ancova_lin,file=paste0(folder,"ancova_",z,"_gymnosperm_lin.RData"))    
  }
  
  fit=fit.ancova_lin
  
  posteriors_fec<-as.data.frame(fit) |>
    dplyr::select(!matches("beta_")) |>
    dplyr::select(matches("beta"))
  colnames(posteriors_fec)[1:ncol(X)]=colnames(X)
  speciescode=sp.zone |> ungroup()|> dplyr::select(species) |> mutate(code=as.numeric(as.factor(species))) |> unique()
  speciesranef=as.data.frame(summary(fit)$summary) |> 
    rownames_to_column(var="parameter") |> 
    filter(grepl("alpha",parameter)) |> 
    dplyr::select(parameter,mean) |> 
    mutate(code=as.numeric(gsub(".*?\\[([^]]*)\\].*", "\\1", parameter))) |> 
    left_join(speciescode) |> 
    dplyr::select(species,mean)
  
  contrast_mat= X |>  t()
  contrast_post=as.matrix(posteriors_fec)[,1:ncol(X)] %*% contrast_mat
  fit=apply(contrast_post,2,median)
  quant05=apply(contrast_post,2,quantile,probs=0.05)
  quant95=apply(contrast_post,2,quantile,probs=0.95)
  
  output=cbind(sp.zone,
               fit=fit,
               quant05=quant05,
               quant95=quant95) |>
    left_join(speciesranef,by="species") 

  out_ancova_taxa=rbind(out_ancova_taxa,
                   output)
  rm(output,quant05,quant95,fit,contrast_post,contrast_mat,speciesranef,
     speciescode,posteriors_fec,data_zone,sp.mod,sp.zone)
}



out_ancova_taxa |>
  mutate(zone=factor(str_to_title(zone),levels=c("West","East","Europe")),
         margin.temp=case_when(margin.temp=="midtemp"~"center",
                               TRUE~margin.temp)) |>
  ggplot()+
  geom_point(aes(dh,log(dISP)-mean,color=margin.temp),size=0.6,alpha=0.05)+
  geom_ribbon(aes(x=dh,ymin=quant05,ymax=quant95,fill=margin.temp),alpha=0.8)+
  geom_line(aes(dh,fit,group=margin.temp),size=0.2,color="black")+
  geom_boxplot(aes(dh,y=-12,color=margin.temp),position=position_dodge(width=2),outlier.colour = NA)+
  scale_fill_manual(values=c("bisque2","steelblue1","firebrick1"))+
  scale_color_manual(values=c("bisque2","steelblue1","firebrick1"))+
  facet_wrap(taxa~zone)+
  theme(text = element_text(size=13),
        legend.position = "bottom")+
  labs(y="Log ratio of fecundity relative to the mean",
       x="Water deficit (mm, centered by species)",
       color="Temperature margin",
       fill="Temperature margin")

# ggsave( p,
#         filename = "bes_fig/Int_2.png",
#         dpi = 600,
#         width = 17,
#         height=13,
#         units = "cm")

```

# Does variation in fecundity in species margin depends on the ecology of species?

## Climatic range

```{r}
fecundity.fit.1 |> 
  group_by(species) |> 
  # compute weighted quantiles of sgdd and wai, and correlation between wai and sgdd
  mutate(mean.dh=mean(dh,na.rm=TRUE),
         mean.mat=mean(mat,na.rm=TRUE),
         cor.mat.dh=cor(dh,mat,use="complete.obs")) |>
  # filter(dh>quantile(dh,probs = 0.01,na.rm=TRUE)) |> filter(dh<quantile(dh,probs=0.99,na.rm=TRUE)) |> 
  filter(dh>(-3000)) |> 
  # filter(!is.na(margin.temp)) |> 
  # filter(!is.na(margin.temp)&!is.na(margin.deficit)) |>
  # ungroup() |> group_by(species,margin.temp) |> 
  mutate(max_mat=quantile(mat,probs=0.99),
         min_mat=quantile(mat,probs = 0.01),
         mean_mat=mean(mat),
         max_dh=quantile(dh,probs=0.99),
         min_dh=quantile(dh,probs = 0.01),
         mean_dh=mean(dh)) |> 
  ungroup() |> 
  # mutate(species=forcats::fct_reorder(species,max),
  #        ISP=fecGmMu/BA,
  #        s_ISP=fecGmSd/(BA^2))|>
  # filter(!is.na(ISP)) |>
  filter(!is.na(dh)) |> 
  ggplot(aes(mean_dh,mean_mat,color=zone))+
  geom_point()+
  # geom_linerange(aes(xmax=max_dh,xmin=min_dh,y=mean_mat,color=zone))+
  # geom_linerange(aes(x=mean_dh,ymax=max_mat,ymin=min_mat,color=zone))
  # geom_smooth(method="lm")+
  theme(legend.position = "bottom")
  # ggplot(aes(max-min,zone))+
  # geom_boxplot()
```

## Model by species

```{r}
tar_load(fitspecies,store="target_fit")
fitspecies |> 
  ungroup() |> 
  # filter(margin.temp=="hot") |> 
  filter(zone=="west") |>
  # sample_n(500) |> 
  # filter(species=="pinusPinea") |> 
  ggplot(aes(dh,fit,color=margin.temp))+
  geom_line()+
  facet_wrap(~species)

plot.sp.margin<-function(mmat,mdh,labmat,labdh){
  plot<-fitspecies |> 
    ungroup() |> 
    group_by(zone,species) |> 
    mutate(margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.05)[[1]]~"humid",
                                dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                   dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                                 dh>weighted.quantile(dh,w=BA, prob=0.95)[[1]]~"arid")) |> 
    filter(margin.temp%in%mmat&margin.deficit%in%mdh)|> #&margin.deficit=="arid"  
    summarise(med.ISP=weighted.quantile(fit,w=1/(quant95-quant05), prob=0.5)[[1]],
              mean.ISP=mean(fit),
              med.dh=median(dh),
              med.mat=median(mat)) |>
    ungroup() |> 
    ggplot()+
    geom_point(aes(med.dh,med.ISP,color=zone))+
    geom_smooth(aes(med.dh,med.ISP,color=zone),method="lm")+
    geom_hline(yintercept = 0,color="darkgrey")+
    geom_vline(xintercept = 0,color="darkgrey")+
    xlim(c(-3000,1500))+
    ylim(c(-6,5))+
    theme(legend.position = "left")+
    labs(title=paste0("Median fecundity at the ",labmat, " and ",labdh," margin"))
  boxplot<-fitspecies |> 
    ungroup() |> 
    group_by(zone,species) |> 
    mutate(margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.05)[[1]]~"humid",
                                dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                   dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                                 dh>weighted.quantile(dh,w=BA, prob=0.95)[[1]]~"arid")) |> 
    filter(margin.temp%in%mmat&margin.deficit%in%mdh)|> #&margin.deficit=="arid"  
    summarise(med.ISP=weighted.quantile(fit,w=1/(quant95-quant05), prob=0.5)[[1]],
              mean.ISP=mean(fit),
              med.dh=median(dh),
              med.mat=median(mat)) |>
    ungroup() |> 
    ggplot(aes(1,med.ISP))+
    geom_boxplot(outlier.colour = NA)+
    geom_hline(yintercept = 0,color="darkgrey")+
    theme(axis.text = element_text( color = "white"),
          axis.title.y = element_blank())+
    labs(x="Isp median",
         title="")+
    ylim(c(-6,5))
  print(cowplot::plot_grid(plot,boxplot, rel_widths = c(3,0.5)))
}

plot.sp.margin("hot","humid","hot","humid")
plot.sp.margin("hot","arid","hot","arid")
plot.sp.margin("cold","humid","cold","humid")
plot.sp.margin("cold","arid","cold","arid")
plot.sp.margin("hot",c("arid","midhum","humid"),"hot","all dh")
plot.sp.margin("cold",c("arid","midhum","humid"),"cold","all dh")
```

```{r}
fitspecies |> 
  filter(zone=="west") |>
  filter(species %in% c("abiesAmabilis","pinusContorta","pinusMonticol","pinusPonderos")) 

fitspecies |> 
  ungroup() |> 
  filter(zone=="west") |>
  mutate(n=n()) |> 
  group_by(species,n) |> 
  summarise(n_sp=n()) |> 
  mutate(perc=100*(n_sp/n)) |> 
  arrange(desc(perc))

```

```{r}
fitspecies |> 
  dplyr::select(-biome) |> 
  left_join(species.biome) |> 
  group_by(species) |> 
  mutate(margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.05)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.95)[[1]]~"arid")) |>
  ungroup() |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot")),
         margin.deficit=factor(margin.deficit,levels=c("humid","midhum","arid")),
         biome=case_when(biome==4~"Temperate deciduous",
                         biome==5~"Temperate coniferous",
                         biome==6~"Boreal",
                         biome==8~"Temperate Grasslands",
                         biome==12~"Mediteranean",
                         biome==13~"Desert")) |>   
  # filter(margin.temp!="midtemp") |>
  filter(!is.na(margin.deficit)) |> 
  ggplot(aes(margin.temp,fit,fill=margin.deficit,color=margin.deficit))+
  geom_boxplot(outlier.colour = NA)+
  geom_hline(yintercept = 0,color="grey")+
  scale_fill_manual(values=c("steelblue1","bisque2","firebrick1"))+
  scale_color_manual(values=c("steelblue1","bisque2","firebrick1"))+
  facet_wrap(~biome, ncol = 3,scales="free")+
  theme(legend.position="bottom")+
  labs(fill="",
       color="",
       x="")


fitspecies |> 
  dplyr::select(-biome) |> 
  left_join(species.biome) |> 
  mutate(margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.05)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.95)[[1]]~"arid")) |>
  ungroup() |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot")),
         margin.deficit=factor(margin.deficit,levels=c("humid","midhum","arid")),
         biome=case_when(biome==4~"Temperate deciduous",
                         biome==5~"Temperate coniferous",
                         biome==6~"Boreal",
                         biome==8~"Temperate Grasslands",
                         biome==12~"Mediteranean",
                         biome==13~"Desert")) |>   
  filter(margin.temp!="midtemp") |>
  filter(!is.na(margin.deficit)) |> 
  mutate(species=forcats::fct_reorder(species,mean.mat)) |> 
  ggplot(aes(margin.deficit,fit,fill=species,color=species))+
  geom_boxplot(outlier.colour = NA)+
  geom_hline(yintercept = 0,color="grey")+
  facet_grid(margin.temp~biome, )+
  theme(legend.position="none")+
  labs(fill="",
       color="",
       x="")

```


## Model by biome 

```{r}
tar_load(fitbiomecontinuous,store="target_fit")
fitbiomecontinuous |> 
  group_by(biome) |> 
  mutate(dh05=quantile(dh,probs=0.05),
         dh425=quantile(dh,probs=0.425),
         dh525=quantile(dh,probs=0.525),
         dh95=quantile(dh,probs=0.95)) |> 
  ggplot()+
  geom_point(aes(dh,log(dISP)-mean,color=margin.temp),size=0.6,alpha=0.05)+
  geom_ribbon(aes(x=dh,ymin=quant05,ymax=quant95,fill=margin.temp),alpha=0.8)+
  geom_line(aes(dh,fit,group=margin.temp),size=0.2,color="black")+
  geom_boxplot(aes(dh,y=-12,color=margin.temp),position=position_dodge(width=2),
               outlier.colour = NA)+
  geom_boxplot(aes(log(dISP)-mean,x=2000,color=margin.temp),position=position_dodge(width=250),outlier.colour = NA)+
  # geom_vline(aes(xintercept=dh05))+
  # geom_vline(aes(xintercept=dh425))+
  # geom_vline(aes(xintercept=dh525))+
  # geom_vline(aes(xintercept=dh95))+
  scale_fill_manual(values=c("bisque2","steelblue1","firebrick1"))+
  scale_color_manual(values=c("bisque2","steelblue1","firebrick1"))+
  facet_wrap(~biome)

fitbiomecontinuous |> 
  group_by(biome,species) |> 
  mutate(margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.05)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.95)[[1]]~"arid")) |> 
  filter(!is.na(margin.deficit)) |> 
  ungroup() |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot")),
         margin.deficit=factor(margin.deficit,levels=c("humid","midhum","arid")),
         biome=case_when(biome==4~"Temperate deciduous",
                         biome==5~"Temperate coniferous",
                         biome==6~"Boreal",
                         biome==8~"Temperate Grasslands",
                         biome==12~"Mediteranean",
                         biome==13~"Desert")) |> 
  # filter(biome!="prairial") |>
  filter(margin.temp!="midtemp") |>
  group_by(biome) |> 
  mutate(n_sp=n_distinct(species)) |> ungroup() |> 
  mutate(biome=paste0(biome," (n species = ",n_sp,")")) |> 
  ggplot(aes(margin.temp,fit,fill=margin.deficit,color=margin.deficit))+
  geom_boxplot(outlier.colour = NA)+
  geom_hline(yintercept = 0,color="grey")+
  scale_fill_manual(values=c("steelblue1","bisque2","firebrick1"))+
  scale_color_manual(values=c("steelblue1","bisque2","firebrick1"))+
  facet_wrap(~biome, ncol = 2)+
  theme(legend.position="bottom")+
  labs(fill="",
       color="",
       x="",
       y="Log ratio of fecundity relative to the species mean")->fig
fig
# ggsave( fig,
#         filename = "BiomeDhTemp.png",
#         dpi = 600,
#         width = 28,
#         height=13,
#         units = "cm")
# anov_biome=lm(log(dISP)~margin.temp*zone+margin.temp*as.factor(biome),data=fec_biome)
# summary(anov_biome)


fitbiomecontinuous |> filter(biome==12) |> 
  filter(dh<(-500)) |> 
  filter(margin.temp=="hot") |> 
  group_by(species) |> 
  summarise(n=n())

fitbiomecontinuous |> filter(biome==12) |>  
  # filter(species=="pinusPinaster") |>
  filter(margin.temp=="hot") |> 
  ggplot()+
  geom_point(aes(dh,log(dISP)-mean,color=species),size=0.6,alpha=0.5)+
  geom_smooth(aes(dh,log(dISP)-mean,color=species),method="lm")
  # geom_ribbon(aes(x=dh,ymin=quant05,ymax=quant95,fill=margin.temp),alpha=0.8)+
  # geom_line(aes(dh,fit,group=margin.temp),size=0.2,color="black")+
  # geom_boxplot(aes(dh,y=-12,color=margin.temp),position=position_dodge(width=2),
  #              outlier.colour = NA)+
  # geom_boxplot(aes(log(dISP)-mean,x=2000,color=margin.temp),position=position_dodge(width=250),outlier.colour = NA)+
  # scale_fill_manual(values=c("bisque2","steelblue1","firebrick1"))+
  # scale_color_manual(values=c("bisque2","steelblue1","firebrick1"))


fitbiomecontinuous |> 
  filter(biome==12) |> 
  filter(margin.temp=="hot") |> 
  ggplot(aes(mat,color=species))+
  geom_density()
```


```{r}
cbind(fec_tot,
      biome=points_biome$BIOME) |> 
  group_by(species) |> 
  # compute weighted quantiles of sgdd and wai, and correlation between wai and sgdd
  mutate(margin.temp=case_when(mat<=weighted.quantile(mat,w=BA, prob=0.1)[[1]]~"cold",
                               mat<weighted.quantile(mat,w=BA, prob=0.55)[[1]]&
                                 mat>weighted.quantile(mat,w=BA, prob=0.45)[[1]]~"midtemp",
                               mat>=weighted.quantile(mat,w=BA, prob=0.9)[[1]]~"hot"),
         margin.temp=factor(margin.temp,levels=c("midtemp","cold","hot"))
  ) |> 
  filter(biome==5) ->test 
test |>
  ggplot(aes(map,mat,color=margin.temp))+geom_point()+
  theme(legend.position = "left")
test |> 
  filter(!is.na(margin.temp)) |> 
  ggplot(aes(dh,color=margin.temp))+
  geom_density()

test |>
  # filter(dh<(-1500)) |>
  filter(block=="america") |> 
  # filter(margin.temp=="hot") |> 
  ggplot()+
  geom_point(aes(lon,lat,color=map))+
  geom_sf(data=worldmap,fill=NA)+
  xlim(c(-130,-80))+
  ylim(c(25,60))+
  scale_color_gradientn(colours = viridis(15))

out_anova_biome |> 
  filter(dh>(-500)) |> 
  filter(biome==5) |> 
  ggplot(aes(dh,log(dISP)-mean,color=margin.temp))+
  geom_point(size=0.6,alpha=0.05)+
  geom_smooth(method = "lm")+
  # geom_ribbon(aes(x=dh,ymin=quant05,ymax=quant95,fill=margin.temp),alpha=0.8)+
  # geom_line(aes(dh,fit,group=margin.temp),size=0.2,color="black")+
  # geom_boxplot(aes(dh,y=-12,color=margin.temp),position=position_dodge(width=2),outlier.colour = NA)+
  # geom_boxplot(aes(log(dISP)-mean,x=2000,color=margin.temp),position=position_dodge(width=250),outlier.colour = NA)+
  # scale_fill_manual(values=c("bisque2","steelblue1","firebrick1"))+
  scale_color_manual(values=c("bisque2","steelblue1","firebrick1"))+
  facet_wrap(~biome)

test |> 
  group_by(plot) |> 
  mutate(ba_tot=sum(BA)) |> 
  filter(margin.temp=="hot") |>
  filter(ba_tot<200) |>
  # filter(dh>500) |> 
  ggplot()+
  geom_point(aes(lon,lat,color=log(ba_tot)))+
  geom_sf(data=worldmap,fill=NA)+
  xlim(c(-130,-80))+
  ylim(c(25,60))+
  scale_color_gradientn(colours = viridis(15))+
  theme(axis.title = element_blank(),
        legend.position="bottom")->a

test |> 
  group_by(plot) |> 
  mutate(ba_tot=sum(BA)) |> 
  filter(margin.temp=="hot") |>
  filter(ba_tot<200) |>
  # filter(dh>500) |> 
  ggplot()+
  geom_point(aes(lon,lat,color=dh))+
  geom_sf(data=worldmap,fill=NA)+
  xlim(c(-130,-80))+
  ylim(c(25,60))+
  scale_color_gradientn(colours = viridis(15))+
  theme(axis.title = element_blank(),
        legend.position="bottom")->b
fig=cowplot::plot_grid(a,b)
# ggsave(fig,
#      filename = "ConiferousBiome.png",
#       dpi = 600,
#       width = 17,
#       height=12,
#       units = "cm")


fec_tot |> 
  group_by(plot) |> 
  summarise(ba_tot=sum(BA))->batot  
  # filter(margin.temp=="hot") |>
  # filter(ba_tot<200) |> 
  # ggplot(aes(ba_tot,fecGmMu))+
  # geom_point()+
  # geom_smooth()
out_anova_biome |> 
  left_join(batot) |> 
  filter(margin.temp=="hot") |>
  filter(ba_tot<200) |> 
  filter(biome==5) |> 
  ggplot(aes(log(ba_tot),log(dISP)-mean,color=margin.temp))+
  geom_point()+
  geom_smooth()

out_anova_biome |> 
  left_join(batot) |> 
  filter(margin.temp=="hot") |>
  filter(ba_tot<200) |> 
  filter(biome%in%c(12,5)) |> 
  sample_n(10000) |>
  mutate(biome=case_when(biome==4~"Temperate deciduous",
                         biome==5~"Temperate coniferous",
                         biome==6~"Boreal",
                         biome==8~"Temperate Grasslands",
                         biome==12~"Mediteranean",
                         biome==13~"Desert")) |> 
  ggplot(aes(dh,log(dISP)-mean,color=log(ba_tot)))+
  geom_point(size=1,alpha=0.5)+
  # geom_ribbon(aes(x=dh,ymin=quant05,ymax=quant95),alpha=0.8)+
  geom_line(aes(dh,fit),size=0.2,color="black")+
  scale_color_gradientn(colours = viridis(15))+
  facet_wrap(biome~zone)+
  labs(y="Log ratio of fecundity relative to the mean",
       x="Water deficit (mm)",
       color="Log competition (m2)")+
  theme(legend.position="bottom")#-> fig

# ggsave( fig,
#         filename = "BiomeCompet.png",
#         dpi = 600,
#         width = 14,
#         height=12,
#         units = "cm")

 out_anova_biome |> 
  group_by(species) |> 
  mutate(margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.05)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.95)[[1]]~"arid")) |> 
  ungroup() |> group_by(biome) |> 
  sample_n(min(5000,n())) |> 
  left_join(batot) |> 
  # filter(margin.temp=="hot") |>
  filter(ba_tot<200) |> 
  filter(biome%in%c(4,5,12)) |> 
  ggplot(aes(log(map),log(ba_tot),color=margin.temp))+
  geom_point(alpha=0.5)+
  geom_smooth(aes(color=NULL),method="loess")+
  facet_wrap(~biome)

fec_tot |> 
  left_join(batot) |> 
  # filter(margin.temp%in%c("hot","cold")) |>
  group_by(species) |> 
  mutate(margin.temp=case_when(mat<=weighted.quantile(mat,w=BA, prob=0.1)[[1]]~"cold",
                               mat<weighted.quantile(mat,w=BA, prob=0.55)[[1]]&
                                 mat>weighted.quantile(mat,w=BA, prob=0.45)[[1]]~"midtemp",
                               mat>=weighted.quantile(mat,w=BA, prob=0.9)[[1]]~"hot"),
         margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.05)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.95)[[1]]~"arid")) |> 
  ungroup() |> 
  filter(ba_tot<200) |>
  left_join(sp_biome) |> 
  filter(biome%in%c(4,5,6,12)) |> 
  filter(margin.deficit%in%c("arid","humid")) |>
  filter(margin.temp%in%c("hot","cold")) |>
  ggplot(aes(margin.deficit,log(ba_tot),color=margin.deficit))+
  geom_boxplot(outlier.colour = NA)+
  geom_signif(comparisons = list(c("arid","humid")),
              map_signif_level = TRUE,
              y_position = c(4, 4))+
  facet_grid(~biome)

fec_biome |> 
  group_by(species) |> 
  # compute weighted quantiles of sgdd and wai, and correlation between wai and sgdd
  mutate(margin.temp=case_when(mat<=weighted.quantile(mat,w=BA, prob=0.1)[[1]]~"cold",
                               mat<weighted.quantile(mat,w=BA, prob=0.55)[[1]]&
                                 mat>weighted.quantile(mat,w=BA, prob=0.45)[[1]]~"midtemp",
                               mat>=weighted.quantile(mat,w=BA, prob=0.9)[[1]]~"hot"),
         margin.temp=factor(margin.temp,levels=c("midtemp","cold","hot"))
  ) |> 
  filter(biome%in%c(4,5,6,8,12,13)) |>
  filter(margin.temp=="hot") |> 
  ggplot(aes(dh,color=as.factor(biome)))+
  geom_density()


fec_tot |> 
  left_join(batot) |> 
  # filter(margin.temp%in%c("hot","cold")) |>
  group_by(species) |> 
  mutate(margin.temp=case_when(mat<=weighted.quantile(mat,w=BA, prob=0.1)[[1]]~"cold",
                               mat<weighted.quantile(mat,w=BA, prob=0.55)[[1]]&
                                 mat>weighted.quantile(mat,w=BA, prob=0.45)[[1]]~"midtemp",
                               mat>=weighted.quantile(mat,w=BA, prob=0.9)[[1]]~"hot"),
         margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.05)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.95)[[1]]~"arid")) |> 
  ungroup() |> 
  filter(ba_tot<200) |>
  left_join(sp_biome) |> 
  filter(biome%in%c(4,5,6,12)) |> 
  select(species,biome,ba_tot,lon,lat,margin.temp,zone,dh,mat) |> 
  filter(margin.temp%in%c("hot","cold")) |> 
  unique() |> 
  ggplot(aes(dh,log(ba_tot),color=margin.temp))+
  geom_point()+
  geom_smooth(method="gam")+
  facet_wrap(~biome)

out_anova_biome |> 
  group_by(biome) |> 
  mutate(margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.005)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.995)[[1]]~"arid")) |> 
  ungroup() |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot")),
         margin.deficit=factor(margin.deficit,levels=c("humid","midhum","arid")),
         biome=case_when(biome==4~"Temperate deciduous",
                         biome==5~"Temperate coniferous",
                         biome==6~"Boreal",
                         biome==8~"Temperate Grasslands",
                         biome==12~"Mediteranean",
                         biome==13~"Desert")) |> 
  filter(margin.temp=="hot") |>
  filter(margin.deficit%in%c("arid","humid")) |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot")),
         margin.deficit=factor(margin.deficit,levels=c("humid","midhum","arid"))) |> 
  left_join(batot) |> 
  group_by(biome,species,margin.deficit) |>
  mutate(med.ba=median(log(ba_tot))) |> ungroup() |> 
  group_by(species) |> 
  mutate(logba=log(ba_tot/mean(ba_tot))) |> 
  ggplot(aes(margin.deficit,log(ba_tot),color=margin.deficit))+
  geom_boxplot(outlier.colour = NA)+
  geom_signif(comparisons = list(c("arid","humid")),
              map_signif_level = TRUE,
              y_position = c(4, 4))+
  facet_wrap(~biome)

```


```{r}
test.lm=test |> 
  filter(!is.na(margin.temp)) |> 
  left_join(batot) |> 
  mutate(ISP=fecGmMu/BA,
         dh=scale(dh),
         ba_tot=scale(log(ba_tot))) |> 
  # filter(margin.temp=="hot") |>
  filter(!is.na(ba_tot)) |> 
  filter(!is.na(log(ISP))) |> 
  group_by(species) |> 
  mutate(ISP=scale(log(ISP)))

summary(nlme::lme(log(ISP) ~ ba_tot * dh, random = ~ 1 | species, data = test.lm))

anova(lm(ISP~ba_tot+dh,data=test.lm))
# rm(test)
lme4::lmer(log(ISP)~ba_tot+ 1|species,data=test.lm)->mF

VarCorr(mF)->mfvar
mfvar
cbind(fec_tot,
      biome=points_biome$BIOME) |> 
  group_by(biome) |> 
  filter(biome%in%c(4,5,6,8,12,13)) |> 
  summarise(cor=cor(dh,mat)) |> 
  ggplot(aes(cor,fill=as.factor(biome)))+
  geom_vline(xintercept = 0)+
  geom_histogram(position="dodge")+facet_wrap(~biome)
```


## case of pinus edulis
weird sub-population wtih low fecundity. Is it spatially distinct?

```{r}
fecundity.fit.1 |> 
  filter(species=="pinusEdulis") |> 
  mutate(ISP=fecGmMu/BA,
         dISP=ISP/median(ISP),
         margin.temp=case_when(mat<=weighted.quantile(mat,w=BA, prob=0.1)[[1]]~"cold",
                               mat<weighted.quantile(mat,w=BA, prob=0.55)[[1]]&
                                 mat>weighted.quantile(mat,w=BA, prob=0.45)[[1]]~"midtemp",
                               mat>=weighted.quantile(mat,w=BA, prob=0.9)[[1]]~"hot"),
         margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot"))
  ) |> 
  filter(!is.na(margin.temp)) |>
  ggplot(aes(log(dISP)))+
  geom_density()

fecundity.fit.1 |> 
  filter(species=="pinusEdulis") |> 
  mutate(ISP=fecGmMu/BA,
         dISP=ISP/median(ISP),
         margin.temp=case_when(mat<=weighted.quantile(mat,w=BA, prob=0.1)[[1]]~"cold",
                               mat<weighted.quantile(mat,w=BA, prob=0.55)[[1]]&
                                 mat>weighted.quantile(mat,w=BA, prob=0.45)[[1]]~"midtemp",
                               mat>=weighted.quantile(mat,w=BA, prob=0.9)[[1]]~"hot"),
         margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot"))
  ) |> 
  mutate(pop=case_when(log(dISP)<(-2.5)~1,
                       TRUE~0)) |> 
  ggplot()+
  geom_sf(data=worldmap,fill=NA)+
  geom_sf(data=ecoregions,aes(fill=ECO_NAME))+
  geom_point(aes(lon,lat,color=pop),alpha=0.5)+
  xlim(c(-115,-100))+
  ylim(c(29,45))+
  scale_color_gradientn(colours = viridis(15))+
  theme(legend.position = "none")


fecundity.fit.1 |> 
  filter(species=="pinusEdulis") |> 
  mutate(ISP=fecGmMu/BA,
         dISP=ISP/median(ISP),
         margin.temp=case_when(mat<=weighted.quantile(mat,w=BA, prob=0.1)[[1]]~"cold",
                               mat<weighted.quantile(mat,w=BA, prob=0.55)[[1]]&
                                 mat>weighted.quantile(mat,w=BA, prob=0.45)[[1]]~"midtemp",
                               mat>=weighted.quantile(mat,w=BA, prob=0.9)[[1]]~"hot"),
         margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot"))
  ) |> 
  mutate(pop=case_when(log(dISP)<(-2.5)~1,
                       TRUE~0)) |> 
  ggplot(aes(mat,margin.deficit,color=as.factor(pop)))+
  geom_boxplot()
```

 
# Does the effect of temperature in margins is conserved among continent for identical genus?

```{r fitgenus}
sp.data=rbind(fecundity.eu_clim |> mutate(block="europe"),
              fecundity.am_clim|> mutate(block="america")) |>
  filter(BA!=0) |> dplyr::select(species) |> unique() |> pull()

genus.com<-phylo.select |> filter(species %in% sp.data) |> dplyr::select(genus,zone) |> unique() |> 
  group_by(genus) |> 
  summarise(nsp=n()) |> 
  filter(nsp==3) |> 
  pull(genus)


sp.genus=rbind(fecundity.eu_clim |> mutate(block="europe"),
                fecundity.am_clim|> mutate(block="america")) |>
  left_join(phylo.select) |> 
  dplyr::select(plot,lon,lat,genus,species,BA,fecGmMu,fecGmSd,sgdd,wai,pet,map,mat,block,zone) |> 
  filter(genus %in% genus.com) |> 
  filter(BA!=0) |> #rm absences
  mutate(dh=12*pet-map) |> 
  group_by(species) |> 
  # compute weighted quantiles of sgdd and wai, and correlation between wai and sgdd
  mutate(margin.temp=case_when(mat<=weighted.quantile(mat,w=BA, prob=0.1)[[1]]~"cold",
                               mat<weighted.quantile(mat,w=BA, prob=0.55)[[1]]&
                                 mat>weighted.quantile(mat,w=BA, prob=0.45)[[1]]~"midtemp",
                               mat>=weighted.quantile(mat,w=BA, prob=0.9)[[1]]~"hot"),
         margin.temp=factor(margin.temp,levels=c("midtemp","cold","hot"))
  ) |>
  filter(dh>quantile(dh,probs = 0.01,na.rm=TRUE)) |> filter(dh<quantile(dh,probs=0.99,na.rm=TRUE)) |> 
  filter(!is.na(margin.temp)) |> 
  # filter(!is.na(margin.temp)&!is.na(margin.deficit)) |>
  ungroup() |> group_by(genus) |> 
  mutate(dh_old=dh,
         dh=scale(dh),
         dh2=dh^2) |> 
  ungroup() |> 
  mutate(ISP=fecGmMu/BA,
         s_ISP=fecGmSd/(BA^2))|> 
  filter(!is.na(ISP)) |> 
  filter(!is.na(dh))

folder="mod_ancova_genus_2/"
out_ancova=sp.genus |> 
  mutate(fit=NA,
         quant05=NA,
         quant95=NA,
         mean=NA) |> 
  slice(0)
for(g in genus.com){
  print(g)
  for (z in c("europe","east","west")){
    print(z)
    # filter data
    sp.mod=sp.genus |> 
      filter(genus==g) |> 
      filter(zone==z)
      
    if(file.exists(paste0(folder,"ancova_",z,"_",g,".RData"))){
      load(paste0(folder,"ancova_",z,"_",g,".RData"))
    }else{
      X=model.matrix(~margin.temp*dh+margin.temp*dh2, sp.mod)
      data_zone=list(N=dim(sp.mod)[1],
                     NX=ncol(X),
                     S=nlevels(as.factor(sp.mod$species)),
                     species=as.numeric(as.factor(sp.mod$species)),
                     ISP=sp.mod$ISP,
                     X=X,
                     NULL)
      fit.ancova <- stan(file = "lmm_ancova.stan",
                         data=data_zone,
                         iter=1000,
                         chains=3,
                         core=3,
                         include=FALSE,
                         pars=c("beta_raw"))
      save(fit.ancova,file=paste0(folder,"ancova_",z,"_",g,".RData"))
    }
    if(file.exists(paste0(folder,"ancova_",z,"_",g,"_lin.RData"))){
      load(paste0(folder,"ancova_",z,"_",g,"_lin.RData"))
    }else{
      X=model.matrix(~margin.temp*dh, sp.mod)
      data_zone=list(N=dim(sp.mod)[1],
                     NX=ncol(X),
                     S=nlevels(as.factor(sp.mod$species)),
                     species=as.numeric(as.factor(sp.mod$species)),
                     ISP=sp.mod$ISP,
                     X=X,
                     NULL)
      fit.ancova_lin <- stan(file = "lmm_ancova.stan",
                             data=data_zone,
                             iter=1000,
                             chains=3,
                             core=3,
                             include=FALSE,
                             pars=c("beta_raw"))
      save(fit.ancova_lin,file=paste0(folder,"ancova_",z,"_",g,"_lin.RData"))    
    }
    
    bic_quad=-2*summary(fit.ancova)$summary["lp__","mean"]+8*dim(sp.mod)[1]
    bic_lin=-2*summary(fit.ancova_lin)$summary["lp__","mean"]+5*dim(sp.mod)[1]
    
    if(bic_quad<bic_lin){
      fit=fit.ancova
      X=model.matrix(~margin.temp*dh+margin.temp*dh2, sp.mod)
    }else{
      fit=fit.ancova_lin
      X=model.matrix(~margin.temp*dh, sp.mod)
    }
    
    posteriors_fec<-as.data.frame(fit.ancova) |>
      dplyr::select(!matches("beta_")) |>
      dplyr::select(matches("beta"))
    colnames(posteriors_fec)[1:ncol(X)]=colnames(X)
    speciescode=sp.mod |> ungroup()|> dplyr::select(species) |> mutate(code=as.numeric(as.factor(species))) |> unique()
    speciesranef=as.data.frame(summary(fit.ancova)$summary) |> 
      rownames_to_column(var="parameter") |> 
      filter(grepl("alpha",parameter)) |> 
      dplyr::select(parameter,mean) |> 
      mutate(code=as.numeric(gsub(".*?\\[([^]]*)\\].*", "\\1", parameter))) |> 
      left_join(speciescode) |> 
      dplyr::select(species,mean)
    
    contrast_mat= X |>  t()
    contrast_post=as.matrix(posteriors_fec)[,1:ncol(X)] %*% contrast_mat
    fit=apply(contrast_post,2,median)
    quant05=apply(contrast_post,2,quantile,probs=0.05)
    quant95=apply(contrast_post,2,quantile,probs=0.95)
    
    output=cbind(sp.mod,
                 fit=fit,
                 quant05=quant05,
                 quant95=quant95) |>
      left_join(speciesranef) 
    
    out_ancova=rbind(out_ancova,
                     output)
    rm(output,quant05,quant95,fit,contrast_post,contrast_mat,speciesranef,speciescode,posteriors_fec,data_zone,sp.mod,sp.zone)
  }
}

out_ancova |> 
  ggplot()+
  geom_line(aes(dh_old,fit,color=margin.temp),size=0.6)+
  geom_point(aes(dh_old,log(ISP)-mean,color=margin.temp),size=0.6,alpha=0.6)+
  geom_ribbon(aes(x=dh_old,ymin=quant05,ymax=quant95,fill=margin.temp),alpha=0.8)+
  scale_fill_manual(values=c("bisque2","steelblue1","firebrick1"))+
  scale_color_manual(values=c("bisque2","steelblue1","firebrick1"))+
  facet_grid(genus~zone)
out_ancova |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot"))) |> 
  ggplot()+
  geom_line(aes(dh_old,fit,color=zone),size=0.6)+
  geom_point(aes(dh_old,log(ISP)-mean,color=zone),size=0.6,alpha=0.1)+
  geom_ribbon(aes(x=dh_old,ymin=quant05,ymax=quant95,fill=zone),alpha=0.8)+
  geom_boxplot(aes(dh_old,y=-10,color=zone),position=position_dodge(width=1),outlier.colour = NA)+
  # geom_boxplot(aes(x=-2000,y=log(ISP)-mean,color=zone),position=position_dodge(width=100))+
  scale_fill_manual(values=c("firebrick1","slateblue","firebrick4"))+
  scale_color_manual(values=c("firebrick1","slateblue","firebrick4"))+
  theme(axis.text.x = element_text(colour = inferno(6)))+
  facet_grid(genus~margin.temp)


```

## Effect of temperature margins only for gymnosperms

```{r}
sp.taxa=rbind(fecundity.eu_clim |> mutate(block="europe"),
                fecundity.am_clim|> mutate(block="america")) |>
  left_join(phylo.select) |> 
  dplyr::select(plot,lon,lat,taxa,genus,species,BA,fecGmMu,fecGmSd,sgdd,wai,pet,map,mat,block,zone) |> 
  filter(taxa == "gymnosperm") |> 
  filter(BA!=0) |> #rm absences
  mutate(dh=12*pet-map) |> 
  group_by(species) |> 
  # compute weighted quantiles of sgdd and wai, and correlation between wai and sgdd
  mutate(margin.temp=case_when(mat<=weighted.quantile(mat,w=BA, prob=0.1)[[1]]~"cold",
                               mat<weighted.quantile(mat,w=BA, prob=0.55)[[1]]&
                                 mat>weighted.quantile(mat,w=BA, prob=0.45)[[1]]~"midtemp",
                               mat>=weighted.quantile(mat,w=BA, prob=0.9)[[1]]~"hot"),
         margin.temp=factor(margin.temp,levels=c("midtemp","cold","hot"))
  ) |>
  filter(dh>quantile(dh,probs = 0.01,na.rm=TRUE)) |> filter(dh<quantile(dh,probs=0.99,na.rm=TRUE)) |>
  filter(!is.na(margin.temp)) |> 
  ungroup() |> 
  mutate(ISP=fecGmMu/BA,
         s_ISP=fecGmSd/(BA^2))|> 
  filter(!is.na(ISP)) |> 
  filter(!is.na(dh))
sp.taxa |> 
  ggplot(aes(margin.temp,ISP))+
  geom_boxplot()+
  scale_y_log10()

mid_isp<-sp.taxa |> 
    group_by(species) |> 
    summarise(mid_isp=median(ISP)[[1]]) |> 
    ungroup() |> 
    dplyr::select(species,mid_isp)
sp.taxa<-sp.taxa |>
  left_join(mid_isp) |>
  mutate(dISP=ISP/mid_isp,
         s_ISP=s_ISP/(mid_isp^2)) |> 
  filter(!is.na(dISP))

X=model.matrix(~margin.temp*dh*zone, sp.taxa)
   
data_gymno=list(N=dim(sp.taxa)[1],
               NX=ncol(X),
               S=nlevels(as.factor(sp.taxa$species)),
               species=as.numeric(as.factor(sp.taxa$species)),
               ISP=sp.taxa$dISP,
               X=X,
               NULL)
fit.ancova_lin <- stan(file = "lmm_ancova.stan",
                   data=data_gymno,
                   iter=1000,
                   chains=3,
                   core=3,
                   include=FALSE,
                   pars=c("beta_raw"))
# save(fit.ancova_lin,file="fit.gymno.margintemp.dh.rdata")
posteriors_fec<-as.data.frame(fit.ancova_lin) |>
  dplyr::select(!matches("beta_")) |>
  dplyr::select(matches("beta"))
colnames(posteriors_fec)[1:ncol(X)]=colnames(X)


speciescode=sp.taxa |> ungroup()|> dplyr::select(species) |> mutate(code=as.numeric(as.factor(species))) |> unique()
speciesranef=as.data.frame(summary(fit.ancova_lin)$summary) |> 
  rownames_to_column(var="parameter") |> 
  filter(grepl("alpha",parameter)) |> 
  dplyr::select(parameter,mean) |> 
  mutate(code=as.numeric(gsub(".*?\\[([^]]*)\\].*", "\\1", parameter))) |> 
  left_join(speciescode) |> 
  dplyr::select(species,mean)

contrast_mat= X |>  t()
contrast_post=as.matrix(posteriors_fec)[,1:ncol(X)] %*% contrast_mat
fit=apply(contrast_post,2,median)
quant05=apply(contrast_post,2,quantile,probs=0.05)
quant95=apply(contrast_post,2,quantile,probs=0.95)

cbind(sp.taxa,
      fit=fit,
      quant05=quant05,
      quant95=quant95) |>
  left_join(speciesranef,by="species") |> 
  mutate(zone=factor(str_to_title(zone),levels=c("West","East","Europe")),
         margin.temp=case_when(margin.temp=="midtemp"~"center",
                               TRUE~margin.temp)) |>
  ggplot()+
  geom_point(aes(dh,log(dISP)-mean,color=margin.temp),size=0.6,alpha=0.05)+
  geom_ribbon(aes(x=dh,ymin=quant05,ymax=quant95,fill=margin.temp),alpha=0.8)+
  geom_line(aes(dh,fit,group=margin.temp),size=0.2,color="black")+
  # geom_boxplot(aes(dh,y=-12,color=margin.temp),position=position_dodge(width=2),outlier.colour = NA)+
  scale_fill_manual(values=c("bisque2","steelblue1","firebrick1"))+
  scale_color_manual(values=c("bisque2","steelblue1","firebrick1"))+
  facet_wrap(~zone)+
  theme(text = element_text(size=13),
        legend.position = "bottom")+
  labs(y="Log ratio of fecundity relative to the mean",
       x="Water deficit (mm, centered by species)",
       color="Temperature margin",
       fill="Temperature margin")

```


# Volatility effect

```{r}
sp.bayes=rbind(fecundity.eu_clim |> mutate(block="europe"),
                 fecundity.am_clim|> mutate(block="america")) |>
  left_join(phylo.select) |> 
  filter(BA!=0) |> #rm absences
  mutate(dh=12*pet-map) |> 
  group_by(species) |> 
  filter(!is.na(fecGmMu)) |> 
  dplyr::select(plot,lon,lat,species,BA,fecGmMuW,fecGmVarW,fecGmCv,volatility,dh,mat,block,zone) |> 
  mutate(margin.temp=case_when(mat<weighted.quantile(mat,w=BA, prob=0.1)[[1]]~"cold",
                               mat<weighted.quantile(mat,w=BA, prob=0.55)[[1]]&
                                 mat>weighted.quantile(mat,w=BA, prob=0.45)[[1]]~"midtemp",
                               mat>weighted.quantile(mat,w=BA, prob=0.9)[[1]]~"hot"),
         margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.1)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.55)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.45)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.9)[[1]]~"arid")) |> 
  mutate(margin.temp=factor(margin.temp,levels=c("midtemp","cold","hot")),
         margin.deficit=factor(margin.deficit,levels=c("midhum","arid","humid"))) |> 
  filter(!is.na(margin.temp)) |>
  ungroup()
sp.bayes |> 
  ggplot(aes(margin.temp,volatility))+ #,color=margin.deficit
  geom_boxplot()+
  scale_y_log10()+
  facet_wrap(~zone)

sp.bayes |> 
  # sample_n(10000) |> 
  ggplot(aes(volatility,fecGmMuW))+
  geom_point()

mid_vol<-sp.bayes |> 
    group_by(species) |> 
    summarise(mid_vol=median(volatility,na.rm=TRUE)[[1]]) |> 
    ungroup() |> 
    dplyr::select(species,mid_vol)
sp.bayes<-sp.bayes |>
  left_join(mid_vol) |>
  mutate(dvol=volatility/mid_vol) |> 
  filter(!is.na(dvol))

sp.bayes |> 
  ggplot(aes(margin.temp,log(dvol),color=zone))+ #,color=margin.deficit
  geom_boxplot(outlier.colour = NA)+
  ylim(c(-3,3))

X=model.matrix(~margin.temp*zone, sp.bayes)
data_list=list(N=dim(sp.bayes)[1],
             NX=ncol(X),
             S=nlevels(as.factor(sp.bayes$species)),
             species=as.numeric(as.factor(sp.bayes$species)),
             ISP=sp.bayes$dvol,
             X=X)
             
fit.lm <- stan(file = "lmm_dif.stan",
                  data=data_list,
                  iter=1000,
                  chains=3,
                  core=3)
# save(fit.lm,file="fit.allsp.volatility.RData")


launch_shinystan(fit.lm)
posteriors_fec<-as.data.frame(fit.lm) |> 
  dplyr::select(!matches("beta_")) |> 
  dplyr::select(matches("beta"))
colnames(posteriors_fec)[1:ncol(X)]=colnames(X)


# contrast.max=
contrast_cor=as.data.frame(X) |> 
  unique() |>
  mutate(margin.temp=case_when(margin.tempcold==1~"cold",
                               margin.temphot==1~"hot",
                               TRUE~"mid"),
         zone=case_when(zoneeurope==1~"europe",
                        zonewest==1~"west",
                        TRUE~"east")) |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","mid","hot"))) |> 
  tibble::rownames_to_column(var = "code") |> 
  dplyr::select(code,margin.temp,zone) |> 
  mutate(contrast=paste0(margin.temp,"_",zone))

contrast_mat= X |> unique() |>  t()
contrast_post=as.matrix(posteriors_fec)[,1:ncol(X)] %*% contrast_mat

as.data.frame(contrast_post) |> 
  pivot_longer(cols=everything(),
               names_to = "code") |>
  left_join(contrast_cor,by="code") |> 
  mutate(zone=factor(zone,levels=c("west","east","europe"))) |> 
  ggplot(aes(margin.temp,value,color=zone))+
  geom_boxplot(outlier.colour = NA)+
  scale_color_manual(values=c("firebrick4","firebrick1","slateblue"))+
  theme(axis.title=element_blank(),
        legend.position = "bottom")+
  geom_hline(yintercept = 0,color="gray12",type="dashed")+
  labs(title="Log ratio of volatility relative to \nthe mean volatility of the niche",
       color="Geographical \nzone:")

```

# figures

## fig1

require output by species and model fitted on biogeographical zones

```{r}
model_data <- fitspecies |>
  ungroup() |> 
  group_by(zone,species) |> 
  mutate(margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.05)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.95)[[1]]~"arid")) |> 
  filter(margin.temp%in%c("cold","hot"))|> # &margin.deficit=="arid" 
  group_by(zone,species,margin.temp) |> 
  summarise(med.ISP=weighted.quantile(fit,w=1/(quant95-quant05), prob=0.5)[[1]],
            mean.ISP=mean(fit),
            med.dh=median(dh),
            med.mat=median(mat)
            ) |>
  ungroup() |> 
  group_by(margin.temp,zone) |> 
  mutate(max.mat=max(med.mat),
         min.mat=min(med.mat)) |> 
  do({
    model = lm(med.ISP ~ med.mat, data = .)
    data.frame(broom::tidy(model), .[1, ])
  }) %>%
  # filter(term == "med.mat") %>%
  group_by(zone,margin.temp) |> 
  mutate(significant = sum(p.value < 0.05)) |> 
  filter(significant>=1) |> 
  dplyr::select(term,estimate,zone,margin.temp,max.mat,min.mat)
model_data=model_data |> 
  pivot_wider(names_from = term,
              values_from = estimate) |> 
  crossing(mat.pred=seq(-10,25,0.1)) |> 
  filter((mat.pred>min.mat)&(mat.pred<max.mat)) |> 
  mutate(pred.isp=`(Intercept)`+mat.pred*med.mat)
```


```{r}
plt<-fitspecies |>
  ungroup() |> 
  group_by(zone,species) |> 
  mutate(margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.05)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.95)[[1]]~"arid")) |> 
  filter(margin.temp%in%c("cold","hot"))|> # &margin.deficit=="arid" 
  group_by(zone,species,margin.temp) |> 
  summarise(med.ISP=weighted.quantile(fit,w=1/(quant95-quant05), prob=0.5)[[1]],
            mean.ISP=mean(fit),
            med.dh=median(dh),
            med.mat=median(mat)) |>
  ungroup() |> 
  ggplot()+
  geom_point(aes(med.mat,med.ISP,color=zone))+
  geom_line(data = model_data,
            aes(mat.pred,pred.isp,color=zone),
            linewidth=1) +
  geom_hline(yintercept = 0,color="gray12")+
  geom_vline(xintercept = 0,color="gray12")+
  # xlim(c(-3000,1500))+
  # ylim(c(-3,3))+ # points out of range
  theme(legend.position = "none",
        axis.text.y = element_text(colour = "white"))+
  scale_color_manual(values=c("firebrick1","slateblue","firebrick4"))+
  labs(title="Species ",
       y="",
       x="Mean annual temperature")+
  facet_wrap(~margin.temp,nrow = 2)
bowp<-fitcontinentdiscrete |>
  filter(margin.temp%in%c("hot","cold")) |>
  ggplot(aes(1,posterior,color=zone))+
  geom_boxplot(outlier.colour = NA)+
  scale_color_manual(values=c("firebrick4","firebrick1","slateblue"))+
  theme(legend.position = "left",
        axis.text.x = element_text(colour = "white"))+
  ylim(c(-3,3))+
  geom_hline(yintercept = 0,color="gray12",type="dashed")+
  labs(x="",
       y="Log ratio of fecundity relative to the mean",
       title="Continent")+
  facet_wrap(~margin.temp,nrow=2)
cowplot::plot_grid(bowp,plt,rel_widths = c(1,2))  -> p
ggsave( p,
        filename = "Respvar.jpg",
        dpi = 600,
        width = 25,
        height=15,
        units = "cm")


```


## fig 2 : structure of climatic space

```{r}
fitspecies |>
  dplyr::select(-biome) |> 
  left_join(species.biome) |>
  ungroup() |> 
  group_by(zone,species,margin.temp) |> 
  summarise(med.ISP=weighted.quantile(fit,w=1/(quant95-quant05), prob=0.5)[[1]],
            med.dh=median(dh),
            med.mat=median(mat)) |>
  filter(margin.temp%in%c("hot","cold")) |> 
  pivot_wider(names_from = margin.temp,values_from = c("med.ISP","med.dh","med.mat")) |> 
  ungroup() |> 
  mutate(disp=med.ISP_hot-med.ISP_cold,
         ddh=med.dh_hot-med.dh_cold,
         dmat=med.mat_hot-med.mat_cold) |> 
  ggplot(aes(zone,ddh,color=as.factor(zone)))+
  geom_boxplot()+
  scale_color_manual(values=c("firebrick1","slateblue","firebrick4"))+
  labs(x="",
       y="Water deficit difference \n (between species hot and cold margins)")+
  theme(legend.position = "none")->ddh

  
fecundity.fit.1 |> 
  dplyr::select(-biome) |> 
  left_join(species.biome) |>
  filter(biome!=8) |> 
  mutate(biome=case_when(biome==4~"Temperate deciduous",
                         biome==5~"Temperate coniferous",
                         biome==6~"Boreal",
                         biome==8~"Temperate Grasslands",
                         biome==12~"Mediteranean",
                         biome==13~"Desert")) |> 
  group_by(biome,zone,species) |> 
  summarise(mean.mat=median(mat),
            mean.dh=median(dh)) |> 
  # summarise(cor=cor(mat,dh, use ="complete" )) |> 
  ggplot(aes(mean.mat,mean.dh,color=biome))+
  geom_jitter(size=0.9,width = 1,height=10)+
  ggalt::geom_encircle(s_shape=0.9,expand=0)+
  scale_color_manual(values=c("deepskyblue","tomato","darkgreen","yellowgreen"))+
  labs(color="Biome",
       x="Species mean MAT (°C)",
       y="Species mean WD (mm)")+
  theme(legend.position="bottom")->matdh
cowplot::plot_grid(matdh,ddh,rel_widths = c(3,1))

fecundity.fit.1 |> 
  dplyr::select(-biome) |> 
  left_join(species.biome) |>
  filter(biome!=8) |> 
  mutate(biome=case_when(biome==4~"Temperate deciduous",
                         biome==5~"Temperate coniferous",
                         biome==6~"Boreal",
                         biome==8~"Temperate Grasslands",
                         biome==12~"Mediteranean",
                         biome==13~"Desert")) |> 
  dplyr::select(zone,biome,species) |> unique() |> 
  ggplot(aes(zone,fill=biome)) +
  geom_bar(stat = "count")+
  scale_fill_manual(values=c("deepskyblue","tomato","darkgreen","yellowgreen"))


fecundity.fit.1 |> 
  dplyr::select(-biome) |> 
  left_join(species.biome) |>
  filter(margin.temp%in%c("cold","hot")) |> 
  group_by(species) |> 
  mutate(cormatdh=cor(mat,dh,use="complete")) |> 
  ungroup() |> 
  dplyr::select(-biome) |> 
  left_join(species.biome) |>
  dplyr::select(zone,biome,species,cormatdh) |> 
  filter(biome!=8) |> 
  mutate(biome=case_when(biome==4~"Temperate deciduous",
                         biome==5~"Temperate coniferous",
                         biome==6~"Boreal",
                         biome==8~"Temperate Grasslands",
                         biome==12~"Mediteranean",
                         biome==13~"Desert")) |> 
  unique() |> 
  ggplot(aes(cormatdh,fill=as.factor(biome)))+
  geom_histogram()+
  facet_wrap(~zone)+
  geom_vline(xintercept = 0,color="darkgrey")+
  # scale_fill_manual(values=c("firebrick1","slateblue","firebrick4"))+
  scale_fill_manual(values=c("deepskyblue","tomato","darkgreen","yellowgreen"))+
  theme(legend.position = "bottom")+
  labs(x="Correlation between temperature and water deficit",
       fill="Biomes",
       # color="Biomes",
       y="")->cordhmat
cowplot::plot_grid(cordhmat,ddh,rel_widths = c(2,1))
```
### all species (even non selected)

```{r}
tar_load(df.nfi.eu,store="target_nfi")
tar_load(df.nfi.am,store="target_nfi")
tar_load(sp.select.eu,store="target_nfi")
tar_load(sp.select.am,store="target_nfi")

load(paste0("data/europe/BA.rdata"))
as.data.frame(BA)|> 
    rownames_to_column(var="plot") |> 
    # dplyr::select(plot,matches(sp.select)) |> 
    pivot_longer(cols=-plot,
                 names_to = "species",
                 values_to = "BA") |> 
  filter(BA!=0)->ba.europe
load(paste0("data/america/BA.rdata"))
as.data.frame(BA)|> 
    rownames_to_column(var="plot") |> 
    # dplyr::select(plot,matches(sp.select)) |> 
    pivot_longer(cols=-plot,
                 names_to = "species",
                 values_to = "BA") |> 
  filter(BA!=0)->ba.america

rbind(df.nfi.am |> mutate(zone="america"),
      df.nfi.eu |> mutate(zone="europe")) |> 
  left_join(rbind(ba.america,
                  ba.europe)) -> nfi.tot

points=sf::st_as_sf(nfi.tot[,c("lon","lat")],coords=c("lon","lat"),crs=sf::st_crs(ecoregions))
points_biome=sf::st_join(points,ecoregions) 

cbind(nfi.tot,
      biome=points_biome$BIOME) |> 
  group_by(species) |> mutate(n_sp=n()) |> 
  group_by(n_sp,species,biome) |>
  summarise(n=n()/n_sp) |> unique() |> ungroup() |> 
  group_by(species) |> 
  arrange(desc(n), .by_group = TRUE) %>%
  slice_head(n = 1) |> 
  filter(n>0.33) |> 
  ungroup() |> 
  select(species,biome,n) |> 
  mutate(biome=case_when(biome==13 ~ 12,
                         TRUE ~ biome))->sp_biome

nfi.tot|> 
  filter(!is.na(BA)) |> 
  mutate(select.sp=species%in%c(sp.select.am,sp.select.eu),
         dh=12*pet-map)|> 
  group_by(select.sp,species) |> 
  mutate(mean.lon=mean(lon),
         zone=case_when(mean.lon > (-10) ~ "europe",
                          mean.lon<(-97)~"west",
                          TRUE~"east")) |> 
  ungroup() |> 
  group_by(zone) |> 
  mutate(corzone=cor(mat,dh,use="complete")) |> 
  ungroup() |> 
  group_by(zone,select.sp) |> 
  mutate(corzone.select=cor(mat,dh,use="complete")) |> 
  ungroup() |> 
  group_by(zone,corzone,corzone.select,select.sp,species) |> 
  summarise(cordhmat=cor(mat,dh,use="complete"))  |>
  ggplot(aes(cordhmat,fill=select.sp))+
  geom_histogram()+
  geom_vline(aes(xintercept=corzone))+
  geom_vline(aes(xintercept=corzone.select,color=select.sp))+
  facet_wrap(~zone)


nfi.tot |> 
  left_join(sp_biome) |> 
  filter(!is.na(BA)) |> 
  mutate(select.sp=species%in%c(sp.select.am,sp.select.eu),
         dh=12*pet-map)|> 
  group_by(select.sp,species) |> 
  mutate(mean.lon=mean(lon),
         zone=case_when(mean.lon > (-10) ~ "europe",
                          mean.lon<(-97)~"west",
                          TRUE~"east")) |> 
  ungroup() |> 
  group_by(biome) |> 
  mutate(corbiome=cor(mat,dh,use="complete")) |> 
  ungroup() |> 
  group_by(biome,select.sp) |> 
  mutate(corbiome.select=cor(mat,dh,use="complete")) |> 
  ungroup() |> 
  group_by(biome,corbiome,corbiome.select,select.sp,species) |> 
  summarise(cordhmat=cor(mat,dh,use="complete"))  |>
  ggplot(aes(cordhmat,color=as.factor(biome)))+
  geom_density(alpha=0.6)

```

## fig3

```{r}
fitbiomecontinuous |> 
  group_by(biome,species) |> 
  mutate(margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.01)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.545)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.495)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.99)[[1]]~"arid")) |> 
  filter(!is.na(margin.deficit)) |> 
  ungroup() |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot")),
         margin.deficit=factor(margin.deficit,levels=c("humid","midhum","arid")),
         biome=case_when(biome==4~"Temperate deciduous",
                         biome==5~"Temperate coniferous",
                         biome==6~"Boreal",
                         biome==8~"Temperate Grasslands",
                         biome==12~"Mediteranean",
                         biome==13~"Desert")) |> 
  # filter(biome!="prairial") |>
  filter(margin.temp!="midtemp") |>
  group_by(biome) |> 
  mutate(n_sp=n_distinct(species)) |> ungroup() |> 
  mutate(biome=paste0(biome," (n species = ",n_sp,")")) |> 
  ggplot(aes(margin.temp,fit,fill=margin.deficit,color=margin.deficit))+
  geom_boxplot(outlier.colour = NA)+
  geom_hline(yintercept = 0,color="grey")+
  scale_fill_manual(values=c("steelblue1","bisque2","firebrick1"))+
  scale_color_manual(values=c("steelblue1","bisque2","firebrick1"))+
  facet_wrap(~biome, ncol = 4)+
  theme(legend.position="top",
        strip.text = element_text(size=12))+
  # theme_bw()+
  labs(fill="",
       color="",
       x="",
       y="Log ratio of fecundity relative to the species mean")->a


fecundity.fit.1 |> 
  group_by(plot) |> 
  summarise(ba_tot=sum(BA))->batot  

fitbiomecontinuous |> 
  group_by(biome) |> 
  mutate(margin.deficit=case_when(dh<weighted.quantile(dh,w=BA, prob=0.005)[[1]]~"humid",
                              dh<weighted.quantile(dh,w=BA, prob=0.525)[[1]]&
                                 dh>weighted.quantile(dh,w=BA, prob=0.475)[[1]]~"midhum",
                               dh>weighted.quantile(dh,w=BA, prob=0.995)[[1]]~"arid")) |> 
  ungroup() |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot")),
         margin.deficit=factor(margin.deficit,levels=c("humid","midhum","arid")),
         biome=case_when(biome==4~"Temperate deciduous",
                         biome==5~"Temperate coniferous",
                         biome==6~"Boreal",
                         biome==8~"Temperate Grasslands",
                         biome==12~"Mediteranean",
                         biome==13~"Desert")) |> 
  filter(margin.temp=="hot") |>
  filter(margin.deficit%in%c("arid","humid")) |> 
  mutate(margin.temp=factor(margin.temp,levels=c("cold","midtemp","hot")),
         margin.deficit=factor(margin.deficit,levels=c("humid","midhum","arid"))) |> 
  left_join(batot) |> 
  group_by(biome,species,margin.deficit) |>
  mutate(med.ba=median(log(ba_tot))) |> ungroup() |> 
  group_by(species) |> 
  mutate(logba=log(ba_tot/mean(ba_tot))) |> 
  ggplot(aes(margin.deficit,log(ba_tot),color=margin.deficit))+
  geom_boxplot(outlier.colour = NA)+
  ggpubr::geom_signif(comparisons = list(c("arid","humid")),
              map_signif_level = TRUE,
              y_position = c(4, 4),
              color="black")+
    scale_color_manual(values=c("steelblue1","firebrick1"))+
  facet_wrap(~biome,ncol=4)+
  theme(legend.position = "none",
        strip.text = element_blank())+
  # theme_tufte()+
  labs(x="",
       y="Total basal area, logged (m2)")->b

cowplot::plot_grid(a,b,rel_heights = c(3,2),nrow = 2)
```


