---
title: "TCGK-22/06/2023"
author: "Anne Baranger"
date: "2023-06-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE,message=FALSE)
library(targets)
lapply(c("stringr","ggplot2","tidyr","dplyr","rworldmap","modi","tibble","terra","factoextra","sf","data.table","viridis","robustbase"),require,character.only=TRUE)
theme_set(theme_minimal())
tar_load(mastif.eu,store="target_data")
tar_load(mastif.am,store="target_data")

#tar_load(meanClimate_species,store="target_gbif")

tar_load(species_selection,store="target_nfi")

df.mastif.tree=rbind(mastif.am$df.tree,mastif.eu$df.tree) 
df.species=rbind(mastif.am$df.species.select,mastif.eu$df.species.select)
sp.select.eu=species_selection$sp.select.eu
sp.select.am=species_selection$sp.select.am


# small format of gbif
df.gbif.short <- species_selection$df.gbif |>
  separate(clim,into = c("clim","range")) |> 
  pivot_wider(names_from = range,
                values_from = clim_niche)

# tar_load(nfi.am_clim,store="target_nfi")
# tar_load(nfi.eu_clim,store="target_nfi")
tar_load(fecundity.eu_clim,store="target_nfi")
tar_load(fecundity.am_clim,store="target_nfi")
```

# Data map 

```{r MastifPlots, fig.cap="Maps of Mastif calibration plots"}

# df.niche=read.csv("output/sp_gbif_climate.csv",sep=" ")
#plot initial dataset and selection
worldmap <- sf::st_as_sf(getMap(resolution = "high"))


# plots of selected species
df.mastif.tree |> 
  filter(species %in% df.species$species) |> 
  select(species,lon,lat) |> 
  distinct() |> 
  ggplot() +
  geom_sf(data=worldmap,fill=NA) +
  geom_point(aes(x=lon,y=lat,color=species))+
  xlim(range(df.species[,c("minlon","maxlon")])) + 
  ylim(range(df.species[,c("minlat","maxlat")]))+
  theme_minimal()+
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "none")+
  labs(color="Species")
```

# Species selection

Table of tree effectives :

```{r PlotsnSp,fig.cap="Numbers of plots per species"}
df.mastif.tree |> 
  filter(!duplicated(treeID)) |> 
  group_by(species) |> 
  mutate(n_ind=n()) |> ungroup() |> 
  select(species,n_ind,plotID) |> 
  unique() |> 
  group_by(species,n_ind) |> 
  summarise(n=n())


```

Species deleted:

```{r spdeleted}
sp.deleted=unique(df.species$species)[!unique(df.species$species)%in% c(sp.select.am,sp.select.eu)] 
for (sp in sp.deleted){
  plot<-species_selection$df.tree_w |>
    select(plotID,species,clim,clim_val,cv_plot,low,high) |> 
    left_join(df.gbif.short,by=c("species","clim")) |> 
    filter(species==sp) |> 
    ggplot()+
    geom_histogram(aes(x=clim_val,y=..density..,weight=1/cv_plot))+
    geom_point(aes(x=opt,y=0),color="darkred",size=1)+
    geom_segment(aes(x=low.y,xend=high.y,y=-0,yend=-0),color="darkred",size=1)+
    geom_vline(aes(xintercept = low.x))+
    geom_vline(aes(xintercept = high.x))+
    facet_wrap(~clim,
               # nrow = 3,
               scales = "free")+
    labs(title=sp)
  print(plot)
  rm(plot)
}
rm(sp.deleted)
```

Species selected: 
```{r}
print(paste0("American species : ",length(sp.select.am)))
print(paste0("European species : ",length(sp.select.eu)))
```
# Fecundity variation with geographic distance to the center - European case


## Variation of ISP with latitude

Not clear.

```{r ISPlat}
fecundity.eu_clim |> 
  filter(BA!=0) |> 
  ggplot(aes(lat,ISP))+
  geom_point()+
  scale_y_log10()+
  geom_smooth(method="gam")

fecundity.eu_clim |> 
  filter(BA!=0) |> 
  ggplot(aes(lat,ISP))+
  geom_point()+
  scale_y_log10()+
  geom_smooth(method="gam")+
  facet_wrap(~species)

fecundity.eu_clim |> 
  filter(BA!=0) |> 
  group_by(species) |> 
  mutate(lat.margin=case_when(lat<quantile(lat,probs=0.05)[1]~"south",
                              lat>quantile(lat,probs=0.95)[1]~"north",
                              lat<quantile(lat,probs=0.525)[1]&
                                lat>quantile(lat,probs=0.475)[1]~"mid")) |> 
  filter(!is.na(lat.margin)) |> 
  ggplot(aes(lat.margin,ISP))+
  geom_boxplot()+
  scale_y_log10()+
  facet_wrap(~species)
```

## Characterize peripheral population with dbscan


```{r ISPdbscan}
# create dataset without absences and get minimum effectives
sp.dataset=as.data.table(fecundity.eu_clim |> filter(BA!=0))
nmin=sp.dataset[BA != 0,.(n = .N),by = species][n == sort(n)[2],n]

# determine centroid
for (sp in unique(sp.dataset$species)){  #c("sorbusAucupari","fagusSylvatic","abiesAlba","piceaAbies")
  print(sp)
  
  # compute centroid of distribution
  sp.geodataset <- st_as_sf(sp.dataset[species==sp,], coords = c("lon", "lat"), crs = 4326)
  sp.centroid <- st_centroid(st_union(sp.geodataset$geometry),ratio=1)
  # plot(sp.geodataset$geometry)
  # plot(sp.centroid,add=TRUE,col="red")
  # n=max(round(dim(sp.dataset[species==sp,])[1]/1000),9)
 
  # create dataset with distance from centroid and sample nmin 
  df.kmeans=cbind(sp.dataset[species==sp,],
      dist=as.numeric(st_distance(sp.geodataset,sp.centroid))) |>
    sample_n(nmin,replace=TRUE)
  #option for sleection only a the furthest points
    # filter(dist>quantile(dist,probs=0.85)[[1]])
    # arrange(desc(dist)) |>
    # slice(1:(max(500,n()/10)))
  
  # print distribution and overlay points selected for margin
  print(
    ggplot()+
      geom_point(data=sp.dataset[species==sp,],aes(x=lon,y=lat))+
      geom_point(data=df.kmeans,aes(x=lon,y=lat),color="red")+
      geom_sf(data=worldmap,fill=NA)+
      xlim(-10,33)+ylim(37,70) # europe
      # xlim(-125,-75)+ylim(24,70) # northamerica
    )
  
  # dbscan
  dbscan::kNNdistplot(df.kmeans[,c("lon","lat")], k =  7)
  eps=quantile(dbscan::kNNdist(df.kmeans[,c("lon","lat")], k =  7),probs=0.85)[[1]]
  print(eps)
  fpc::dbscan(df.kmeans[,c("lon","lat")],eps=0.5,MinPts =7)->df.dbscan

  # print dbscan clusters
  print(
    cbind(df.kmeans,
          cluster=df.dbscan$cluster) |>
      filter(cluster!=0) |> 
      ggplot()+
      geom_sf(data=worldmap,fill=NA,alpha=0.4)+
      geom_point(aes(x=lon,y=lat,colour=as.factor(cluster)))+
      xlim(-10,33)+ylim(37,70) # europe
      # xlim(-125,-75)+ylim(24,70) # northamerica
  )
  
  
  # boxplots of fecundity in clusters with enough individuals
  print(
    cbind(df.kmeans,
          cluster=as.character(df.dbscan$cluster))|> 
      filter(cluster !=0) |> 
      group_by(cluster) |> 
      mutate(n.cluster=n(),
             mean.lat=mean(lat),
             mean.dist=mean(dist)) |> 
      ungroup() |>
      filter(n.cluster>40) |> 
      mutate(order.min=dense_rank(mean.dist)) |> 
      filter(order.min==1 | order.min > max(order.min)-3) |> 
      # filter(cluster=="core"|
      #          mean.lat==max(mean.lat)|
      #          mean.lat==min(mean.lat)) |> 
      mutate(cluster=forcats::fct_reorder(as.factor(cluster),mean.dist)) |> 
      ggplot(aes(cluster,ISP,fill=mean.lat))+
      geom_boxplot()+
      scale_y_log10()+
      ggtitle(label=sp)

  )
  
   print(
    cbind(df.kmeans,
          cluster=as.character(df.dbscan$cluster))|> 
      filter(cluster !=0) |> 
      ggplot(aes(dist,ISP,fill=lat))+
      geom_point()+
      scale_y_log10()+
      geom_smooth(method="gam")+
      ggtitle(label=sp)

  )
  }
  # dbscan::kNNdistplot(df.kmeans[,c("lon","lat")], k =  7)
  # 
  # # boxplots of fecundity in clusters with enough individuals
  # print(
  #   bind_rows(
  #     # core population
  #     cbind(sp.dataset[species==sp,],
  #           dist=as.numeric(st_distance(sp.geodataset,sp.centroid))) |> 
  #       arrange(dist) |>
  #       slice(1:(max(300,n()/15))) |> 
  #       mutate(cluster="core"),
  #     # clusters
  #     cbind(df.kmeans,
  #       cluster=as.character(df.dbscan$cluster))
  #   ) |> 
  #     filter(cluster !=0) |> 
  #     group_by(cluster) |> 
  #     mutate(n.cluster=n(),
  #            mean.lat=mean(lat),
  #            mean.dist=mean(dist)) |> 
  #     ungroup() |>
  #     filter(n.cluster>40) |> 
  #     mutate(order.min=dense_rank(mean.dist)) |> 
  #     filter(order.min==1 | order.min > max(order.min)-3) |> 
  #     # filter(cluster=="core"|
  #     #          mean.lat==max(mean.lat)|
  #     #          mean.lat==min(mean.lat)) |> 
  #     mutate(cluster=forcats::fct_reorder(as.factor(cluster),mean.dist)) |> 
  #     ggplot(aes(cluster,ISP,fill=mean.lat))+
  #     geom_boxplot()+
  #     scale_y_log10()
  # 
  # )

#   # code with k-means algo
#   # kmeans
#   kmeans(df.kmeans[,c("lon","lat")],4)->sum.kmeans
#   ## map clusters
#   print(cbind(df.kmeans,
#         cluster=sum.kmeans$cluster) |> 
#     ggplot(aes(x=lon,y=lat,colour=as.factor(cluster)))+
#     geom_point())
#   ## draw boxplots
#   print(cbind(df.kmeans,
#         cluster=sum.kmeans$cluster) |> 
#     group_by(cluster) |> 
#     mutate(m.dist=mean(dist)) |> 
#     ungroup() |> 
#     mutate(order.min=dense_rank(m.dist)) |> 
#     filter(order.min==1 | order.min > max(order.min)-3) |> 
#     mutate(cluster=forcats::fct_reorder(as.factor(cluster),m.dist)) |> 
#     ggplot(aes(as.factor(cluster),ISP,fill=m.dist))+
#     geom_boxplot()+
#     scale_y_log10())

rm(sp.dataset,nmin,sp.geodataset,sp.centroid,df.kmeans,eps)


```
## Kmeans

```{r ISPdistgeo}
# create dataset without absences and get minimum effectives
sp.dataset=as.data.table(fecundity.eu_clim |> filter(BA!=0))
nmin=sp.dataset[BA != 0,.(n = .N),by = species][n == sort(n)[2],n]

# determine centroid
for (sp in unique(sp.dataset$species)){  #c("sorbusAucupari","fagusSylvatic","abiesAlba","piceaAbies")
  print(sp)
  
  # compute centroid of distribution
  sp.geodataset <- st_as_sf(sp.dataset[species==sp,], coords = c("lon", "lat"), crs = 4326)
  sp.centroid <- st_centroid(st_union(sp.geodataset$geometry),ratio=1)
  # plot(sp.geodataset$geometry)
  # plot(sp.centroid,add=TRUE,col="red")
  # n=max(round(dim(sp.dataset[species==sp,])[1]/1000),9)
 
  # create dataset with distance from centroid and sample nmin 
  df.kmeans=cbind(sp.dataset[species==sp,],
      dist=as.numeric(st_distance(sp.geodataset,sp.centroid))) |>
    sample_n(nmin,replace=TRUE)
  #option for sleection only a the furthest points
    # filter(dist>quantile(dist,probs=0.85)[[1]])
    # arrange(desc(dist)) |>
    # slice(1:(max(500,n()/10)))


  # code with k-means algo
  # kmeans
  kmeans(df.kmeans[,c("lon","lat")],15)->sum.kmeans
  ## map clusters
  print(cbind(df.kmeans,
        cluster=sum.kmeans$cluster) |>
    ggplot()+
    geom_point(aes(x=lon,y=lat,color=as.factor(cluster)))+
    geom_sf(data=worldmap,fill=NA)+
    xlim(-10,33)+ylim(37,70) # europe
      # xlim(-125,-75)+ylim(24,70) # northamerica
    )
  ## draw boxplots
  print(cbind(df.kmeans,
        cluster=sum.kmeans$cluster) |>
    group_by(cluster) |>
    mutate(n.cluster=n(),
           mean.lat=mean(lat),
           mean.dist=mean(dist)) |> 
    ungroup() |>
    mutate(order.min=dense_rank(mean.dist)) |>
    filter(order.min==1 | order.min > max(order.min)-3) |>
    mutate(cluster=forcats::fct_reorder(as.factor(cluster),mean.dist)) |>
    ggplot(aes(as.factor(cluster),ISP,fill=mean.dist))+
    geom_boxplot()+
    scale_y_log10()+
    ggtitle(label=sp)
    )
  }



rm(sp.dataset,nmin,sp.geodataset,sp.centroid,df.kmeans,eps)


```

## Continuous comparison of ISP with geographic distance and latitude
Tentative de LM et RQ.
```{r}
sp.dataset=as.data.table(fecundity.eu_clim |> filter(BA!=0))


# determine centroid
sp.dist=sp.dataset |> 
  st_as_sf(coords = c("lon", "lat"), crs = 4326) |> 
  group_by(species) |> 
  mutate(centroid=st_centroid(st_union(geometry))) |> 
  mutate(lat=st_coordinates(geometry)[,2],
         x_centroid=st_coordinates(centroid)[,1],
         y_centroid=st_coordinates(centroid)[,2]) |> 
  ungroup() |> 
  mutate(dist=as.numeric(st_distance(geometry,centroid,by_element = TRUE)))

sp.dist |> 
  group_by(species) |> 
  mutate(ISP=scale(ISP),
         dist=scale(dist)) |> 
  mutate(dist=cut(dist,
                  breaks = 10)) |> 
  ggplot(aes(dist,ISP))+
  geom_boxplot()+
  scale_y_log10()+
  facet_wrap(~species,scales="free")


# lat/dist plots
sp.dist |> 
  group_by(species) |> 
  mutate(scale=ISP) |>  
  mutate(dist=cut(dist,breaks=10),
         lat=cut(lat,breaks=10)) |> ungroup() |>
  group_by(species,dist,lat) |> 
  summarise(mean.fec=mean(ISP)) |> 
  ggplot(aes(x=dist,y=lat))+
  geom_tile(aes(fill=log(mean.fec)))+
  scale_fill_distiller(palette="YlGnBu",direction=1)+
  facet_wrap(~species,scales="free")
rm(sp.dist,sp.dataset)
```

# Fecundity variability with climatic distance to the center 

## Using margins defined with PCA first axis

### ACP

```{r ACPclim}
# Plot the variables in the first PCA pane for America
tar_load(c(nfi.eu_clim,nfi.am_clim),store="target_nfi")
fviz_pca_var(nfi.am_clim$clim_pca, col.var = "contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, ggtheme = theme_classic())

# Plot the variables in the first PCA pane for Europe
fviz_pca_var(nfi.eu_clim$clim_pca, col.var = "contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, ggtheme = theme_classic())
rm(nfi.eu_clim,nfi.am_clim)

```

### Weighted distribution
```{r}
# # graph densité fécondité corrigée a chaque marge 
# fecundity.eu_clim |> 
#   filter(!is.na(margin)) |> 
#   filter(ISP!=0) |>
#   ggplot()+
#   # geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
#   geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
#   facet_wrap(~species,scales="free")+
#   scale_x_log10()+
#   theme_minimal()
# 
# # graph densité fécondité non corrigée a chaque marge 
# fecundity.eu_clim |> 
#   filter(!is.na(margin)) |> 
#   filter(ISP!=0) |>
#   ggplot()+
#   geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
#   # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
#   facet_wrap(~species,scales="free")+
#   scale_y_log10()+
#   theme_minimal()

fecundity.eu_clim |> 
  filter(!is.na(margin)) |> 
  filter(ISP!=0) |>
  ggplot()+
  geom_boxplot(aes(margin,ISP,color=margin),outlier.colour = NA)+
  # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_y_log10()+
  theme_minimal()+
  ggtitle(label = "Europe")

# graph densité fécondité corrigée a chaque marge 
fecundity.am_clim |> 
  filter(!is.na(margin)) |> 
  filter(ISP!=0) |>
  ggplot()+
  # geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
  geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_x_log10()+
  theme_minimal()+
  ggtitle(label="America")
```

## Delineate margin per species

```{r ACPsp}
clim_var=c("wai","sgdd")
# sp="abiesAlba"
sp="piceaAbies"
for (sp in unique(fecundity.eu_clim$species)){
  sp.dataset=as.data.table(fecundity.eu_clim |> filter(BA!=0)) |> select(-PC1,-PC2)

  print(sp)
  sp.dataset=sp.dataset[species==sp,]
  clim_pca<-prcomp(sp.dataset[,..clim_var] |> drop_na(),
                     scale. = TRUE) 
  print(fviz_pca_biplot(clim_pca,geom=c("point")))
  sp.pca <- cbind(sp.dataset,
                  predict(clim_pca, newdata = sp.dataset[,..clim_var])) |> 
     mutate(margin.sp=case_when(PC1>quantile(PC1,probs=0.95,na.rm=TRUE)[[1]]~"up",
                                PC1<quantile(PC1,probs=0.05,na.rm=TRUE)[[1]]~"low",
                                PC1<quantile(PC1,probs=0.525,na.rm=TRUE)[[1]]&
                                  PC1>quantile(PC1,probs=0.475,na.rm=TRUE)[[1]]~"mid"))
  # fviz_pca_ind(clim_pca,geom.ind="point")
  # km.pca=kmeans(sp.pca[,c("PC1","PC2")], centers = 10)
  # cbind(sp.pca,
  #       cluster=km.pca$cluster)|> 
  #   ggplot(aes(PC1,PC2,color=cluster))+geom_point()
  # cbind(sp.pca,
  #       cluster=km.pca$cluster)|> 
  #   ggplot(aes(PC1,PC2,color=cluster))+geom_point()

  # cbind(sp.pca,
  #       cluster=km.pca$cluster)|> 
  #   mutate(dist=sqrt(PC1^2+PC2^2)) |> 
  #   group_by(cluster) |> 
  #   mutate(mean.dist=mean(dist)) |> ungroup() |> 
  #   mutate(cluster=forcats::fct_reorder(as.factor(cluster),mean.dist)) |> 
  #    ggplot()+
  #   geom_boxplot(aes(as.factor(cluster),ISP,weight=1/cv_plot))+
  #   theme_minimal()+
  #   scale_y_log10()
  print(
    sp.pca|>
      ggplot()+
      geom_point(aes(x=lon,y=lat,color=margin.sp))+
      geom_sf(data=worldmap,fill=NA)+
      xlim(-10,33)+ylim(37,70)
  )
  print(sp.pca |> 
    pivot_longer(cols=c("margin","margin.sp"),
                 names_to="margin",
                 values_to = "type") |> 
    filter(!is.na(type)) |> 
    ggplot()+
    geom_boxplot(aes(type,ISP,weight=1/cv_plot))+
    theme_minimal()+
    scale_y_log10()+
    ggtitle(label=sp)+
    facet_wrap(~margin,scale="free_x"))
}

rm(sp.dataset,clim_pca,sp.pca)
```


## Continous distance - Mahanolinis 

```{r ClimMahanobilis}
# sp="abiesAlba"
for(sp in unique(fecundity.eu_clim$species)){
  fec.sp=fecundity.eu_clim |> 
    filter(species==sp) |> 
    filter(!is.na(BA)) |> 
    mutate(across(c("mat","tmin","map","sgdd","pet","wai"),
                  scale))

  # geographic centroid and distance
  points_sp<-st_union(st_as_sf(fec.sp[,c("lon","lat")],coords=c("lon","lat")))
  centroid_sp<-st_centroid(st_convex_hull(points_sp))
  # centroid_sp2<-st_centroid(points_sp)
  fec.sp=cbind(fec.sp,
               dist_geo=st_distance(st_as_sf(fec.sp[,c("lon","lat")],
                                             coords=c("lon","lat")),
                                    centroid_sp))
  
  print(
    fec.sp |> 
      ggplot(aes(log(ISP),dist_geo))+
      geom_point(alpha=0.4,stroke=NA)+
      geom_smooth(method="lm")+
      ggtitle(label = paste0("geo_",sp))
  )
  
  
  # estimate climate centroid and distance
  fec.sp<-fec.sp[,c("lon","lat","margin","dist_geo","BA","ISP","mat","wai","tmin","sgdd")] |> 
    filter(!is.na(mat)&!is.na(wai)&!is.na(tmin)&!is.na(sgdd))
  sp.elp=covMcd(fec.sp[,c("mat","wai","tmin","sgdd")],alpha=0.7)
  sp.center=c(as.matrix(sp.elp$center))
  
  fec.sp=cbind(fec.sp,
               dist_maha=mahalanobis(fec.sp[,c("mat","wai","tmin","sgdd")],
                                     sp.center,
                                     cov(fec.sp[,c("mat","wai","tmin","sgdd")] |> drop_na())))
  
  
  
  print(
    fec.sp |> 
      ggplot(aes(log(ISP),dist_maha))+
      geom_point(alpha=0.4,stroke=NA)+
      geom_smooth(method="lm")+
      ggtitle(label = paste0("clim_",sp))
  )
    
  print(
    fec.sp |>
      ggplot(aes(margin,dist_maha))+
      geom_boxplot()
  )

}
# fec.sp |> 
#   ggplot(aes(dist_geo,dist_maha))+geom_point()+geom_smooth()
# fec.sp |> 
#   ggplot(aes(margin,dist_maha))+geom_boxplot()
# fec.sp |> 
#   ggplot(aes(lat,log(BA)))+geom_point()+geom_smooth(method="lm")

```


# Fecundity variability in negative safety margins

```{r sfm}

# load data from first paper
psi_min=rast(fread("../../Era5-land/output/psihorday_real.csv")[,c("x","y","psi")],crs="epsg:4326")
frost.index.winter=min(rast("../../Era5-land/data/CHELSA/CHELSA_EUR11_tasmin_month_min_19802005.nc"),na.rm=FALSE)
species.traits=readRDS("../../Era5-land/target_safetymargin/objects//df.traits") |> 
  mutate(sp8=paste0(substr(tolower(genus),1,8),substr(str_to_title(species),1,8))) 
mod.output=readRDS("../../Era5-land/target_analysis/objects/df.mod.select") |> 
  separate(species.binomial, into=c("genus","sp")) |> 
  mutate(species=paste0(substr(tolower(genus),1,8),substr(str_to_title(sp),1,8)),
         t_hsm=na_if(t_hsm,0),
         t_fsm=na_if(t_fsm,0)) 

# extract safety margins for european nfi and correct it with model uptut
fec.eu_sfm=cbind(fecundity.eu_clim,
                 psi=extract(psi_min,data.frame(x=fecundity.eu_clim$lon,
                                      y=fecundity.eu_clim$lat))[[2]],
                 frost=extract(frost.index.winter,data.frame(x=fecundity.eu_clim$lon,
                                      y=fecundity.eu_clim$lat))[[2]]) |> 
  left_join(species.traits[,c("sp8","px.mu","lt50.mean")],by=c("species"="sp8")) |> 
  left_join(mod.output[,c("t_fsm","t_hsm","species","mod")],by="species") |> 
  mutate(hsm=psi-t_hsm*1000-px.mu*1000,
         fsm=frost-t_fsm-lt50.mean,
         margin.sm=case_when(mod%in%c("2sm","hsm")&hsm<0~"drought",
                             mod%in%c("2sm","fsm")&fsm<0~"frost",
                             mod%in%c("2sm")&hsm<0&fsm<0~"both",
                             mod%in%c("2sm")&hsm>0&fsm>0~"core",
                             mod%in%c("hsm")&hsm>0~"core",
                             mod%in%c("fsm")&fsm>0~"core"))

# graph densité fécondité corrigée a chaque marge 
fec.eu_sfm |> 
  filter(!is.na(margin.sm)) |> 
  filter(ISP!=0) |>
  ggplot()+
  # geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
  geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin.sm))+
  facet_wrap(~species,scales="free")+
  scale_x_log10()+
  theme_minimal()+
  ggtitle(label="Fecundity in negative safety margins")

# redefine core
fec.eu_sfm |> 
  filter(!is.na(margin.sm)) |> 
  filter(ISP!=0) |>
  group_by(species) |> 
  mutate(hsm.core=hsm>quantile(hsm,probs=0.8,na.rm=TRUE),
         fsm.core=fsm>quantile(fsm,probs=0.8,na.rm=TRUE)) |> 
  filter(!(margin.sm=="core"&hsm.core==FALSE&fsm.core==FALSE)) |> 
  ggplot()+
  # geom_boxplot(aes(margin,ISP,weight=cv_plot,color=margin),outlier.colour = NA)+
  geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin.sm))+
  facet_wrap(~species,scales="free")+
  scale_x_log10()+
  theme_minimal()+
  ggtitle(label="Fecundity in negative safety margins, tuned core")


# graph densité fécondité non corrigée a chaque marge 
fec.eu_sfm |> 
  filter(!is.na(margin.sm)) |> 
  filter(ISP!=0) |>
  ggplot()+
  geom_boxplot(aes(margin.sm,ISP,color=margin.sm),outlier.colour = NA)+
  # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_y_log10()+
  theme_minimal()


fec.eu_sfm |> 
  filter(!is.na(margin.sm)) |> 
  filter(ISP!=0) |> 
  group_by(species,margin.sm) |> summarise(n=n())
rm(frost.index.winter,psi_min,mod.output,species.traits,fec.eu_sfm)
```

# Temporal variability of fecundity
Petite hésitation sur la méthode de calcul
```{r temporalvar}
# load fecGm time serie
files=list.files("data/europe/")[grepl("fecGmMu-",list.files("data/europe/"))]


df.temporal= fecundity.eu_clim |> 
  select(plot,species,BA,fecGmMu,fecGmSd,lon,lat,CV,mat,tmin,map,sgdd,margin)

for (f in files){
  load(paste0("data/europe/",f))
  year=substr(f,9,12)
  mu=summary |> 
    as.data.frame() |> 
    rownames_to_column(var="plot") |> 
    select(plot,matches(unique(fecundity.eu_clim$species))) |> 
    pivot_longer(cols=-plot,
                 names_to = "species",
                 values_to = paste0("fecGmMu_",year))
  rm(summary)
  load(paste0("data/europe/fecGmSd-",year,".rdata"))
  sd=summary |> 
    as.data.frame() |> 
    rownames_to_column(var="plot") |> 
    select(plot,matches(unique(fecundity.eu_clim$species))) |> 
    pivot_longer(cols=-plot,
                 names_to = "species",
                 values_to = paste0("fecGmSd_",year))
  df.temporal <- df.temporal |> 
    left_join(cbind(mu,sd[3]),by=c("plot","species"))
  rm(summary,mu,sd,year)
}

# transform to long format
df.temporal2 <- df.temporal|> 
  filter(!is.na(margin)) |> 
  pivot_longer(cols=matches("fecGmMu_"),
               names_to=c("var","year"),
               names_sep = "_",
               values_to="fecYear") |> 
  pivot_longer(cols=matches("fecGmSd_"),
               names_to=c("var2","year2"),
               names_sep = "_",
               values_to="fecSdYear") |> 
  filter(year==year2) |> select(-year2)


# graph of CV with climatic margin
df.temporal2 |> 
  mutate(fecYear=na_if(fecYear,0),
         fecSdYear=na_if(fecSdYear,0),
         feccvyear=fecYear/fecSdYear) |> 
  filter(!is.na(fecYear)) |> filter(!is.na(fecSdYear)) |> 
  group_by(margin,plot,species) |> 
  summarise(n_plot=n(),
         fecGmMu_plot=weighted.mean(fecYear,w=1/feccvyear,na.rm=TRUE), # perform weighted mean of the tree on the plot
         se_plot=sqrt(sum(fecSdYear^2)/n_plot),
         cv_plot=se_plot/fecGmMu_plot) |> 
  ungroup() |> 
  ggplot()+
  geom_boxplot(aes(margin,cv_plot,color=margin),outlier.colour = NA)+
  # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_y_log10()+
  theme_minimal()
df.temporal2 |> 
  mutate(fecYear=na_if(fecYear,0),
         fecSdYear=na_if(fecSdYear,0),
         feccvyear=fecYear/fecSdYear) |> 
  filter(!is.na(fecYear)) |> filter(!is.na(fecSdYear)) |> 
  group_by(margin,plot,species) |> 
  summarise(n_plot=n(),
         fecGmMu_plot=weighted.mean(fecYear,w=1/feccvyear,na.rm=TRUE), # perform weighted mean of the tree on the plot
         se_plot2=sqrt(sum(fecSdYear^2)/n_plot),
         se_plot=sqrt(sum((1/feccvyear)*(fecYear-mean(fecYear,na.rm=TRUE)))/sum(1/feccvyear)),
         cv_plot=se_plot/fecGmMu_plot) |> 
  ungroup() |> 
  ggplot()+
  geom_boxplot(aes(margin,cv_plot,color=margin),outlier.colour = NA)+
  # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_y_log10()+
  theme_minimal()
# graph of fecundity in differnet margin (for comparison)
df.temporal2 |> 
  mutate(fecYear=na_if(fecYear,0),
         fecSdYear=na_if(fecSdYear,0),
         feccvyear=fecYear/fecSdYear) |> 
  filter(!is.na(fecYear)) |> filter(!is.na(fecSdYear)) |> 
  group_by(margin,plot,species) |> 
  summarise(n_plot=n(),
         fecGmMu_plot=weighted.mean(fecYear,w=1/feccvyear,na.rm=TRUE), # perform weighted mean of the tree on the plot
         se_plot=sqrt(sum(fecSdYear^2)/n_plot),
         cv_plot=se_plot/fecGmMu_plot) |> 
  ungroup() |> 
  ggplot()+
  geom_boxplot(aes(margin,fecGmMu_plot,weight=cv_plot,color=margin),outlier.colour = NA)+
  # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
  facet_wrap(~species,scales="free")+
  scale_y_log10()+
  theme_minimal()

# both
# df.temporal2 |> 
#   mutate(fecYear=na_if(fecYear,0),
#          fecSdYear=na_if(fecSdYear,0),
#          feccvyear=fecYear/fecSdYear) |> 
#   filter(!is.na(fecYear)) |> filter(!is.na(fecSdYear)) |> 
#   group_by(margin,plot,species) |> 
#   summarise(n_plot=n(),
#          fecGmMu_plot=weighted.mean(fecYear,w=1/feccvyear,na.rm=TRUE), # perform weighted mean of the tree on the plot
#          se_plot=sqrt(sum(fecSdYear^2)/n_plot),
#          cv_plot=se_plot/fecGmMu_plot) |> 
#   ungroup() |> 
#   pivot_longer(cols=c("fecGmMu_plot","cv_plot"),
#              names_to="var",
#              values_to = "val") |> 
#   group_by(species,var) |> mutate(val=scale(val)) |> 
#   ggplot()+
#   geom_boxplot(aes(margin,val,color=var),outlier.colour = NA)+
#   # geom_density(aes(x=ISP,y=..density..,weight=cv_plot,color=margin))+
#   facet_wrap(~species,scales="free")+
#   scale_y_log10()+
#   theme_minimal()

rm(files,df.temporal,df.temporal2)
```

